---
title: "dev_block_spp_list"
author: "N.M. Tarr"
date: "2023-10-06"
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Purpose
This document details and demonstrates a function that retrieves a list of reported species from a particular atlas block.

```{r}
block_spp_lists
```

# Usage

Get summer species lists for a block of interest (May 15 - August 15).
```{r}
block <- "35078H8SE"
spp_lists <- block_spp_lists(block = block, start_day = 135, end_day = 227,
                             within = TRUE)
```

Combine the lists to get a list of all species reported.
```{r}
summer_spp_observed <- spp_lists$all 
```


Get a list of species predicted to occur.
```{r}
source <- "GAP"
season <- "summer"

season <- str_to_upper(season)
source_lookup <- list("GAP" = "GAP_SPP", "eBird" = "EBD_SPP")
source <- source_lookup[[source]]

# Connect to the blocks collection (table)
connection_blocks <- connect_ncba_db(database = "ebd_mgmt", 
                                     collection = "blocks")

# Define and execute a query (with fields) for blocks of predicted presence.
fields <- str_interp('{"${source}": true}')
query <- str_interp('{"ID_BLOCK_CODE" : "${block}"}')
pres <- connection_blocks$find(query = query, fields = fields) %>%
  unnest(source)

# GAP
if (source == "GAP_SPP") {
  # Summer list
  summer_spp <- pres %>% 
    filter(SUMMER == 1)
  summer_spp <- summer_spp$PRIMARY_COM_NAME
  
  # Winter list
  winter_spp <- pres %>% 
    filter(WINTER == 1)
  winter_spp <- winter_spp$PRIMARY_COM_NAME  
} 

if (source == "EBD_SPP") {
  # Summer list
  summer_spp <- pres %>% 
    filter(BREEDING == 1)
  summer_spp <- summer_spp$PRIMARY_COM_NAME
  
  # Winter list
  winter_spp <- pres %>% 
    filter(WINTERING == 1)
  winter_spp <- winter_spp$PRIMARY_COM_NAME
}
```

Which species were predicted to occur during summer but have not been observed?
```{r}
print(setdiff(summer_spp, spp_lists$all))
```

Which species were not predicted to occur but were then observed?
```{r}
print(setdiff(spp_lists$all, summer_spp))
```


Which species have been observed, but not coded as probable or confirmed during summer?
```{r}
print(setdiff(spp_lists$all, union(spp_lists$probable, spp_lists$confirmed)))
```

Species count table
```{r}
# Make empty data frame
dataframe <- data.frame(
  highest_category = c("observed", "possible", "probable", "confirmed"),
  species_count = c(0, 0, 0, 0),
  percent_observed = c(0, 0, 0, 0),
  percent_predicted = c(0, 0, 0, 0)
)

# Fill out values
dataframe[dataframe$highest_category == "observed", 
          "species_count"] <- length(spp_lists$observed_highest)
dataframe[dataframe$highest_category == "possible", 
          "species_count"] <- length(spp_lists$possible_highest)
dataframe[dataframe$highest_category == "probable", 
          "species_count"] <- length(spp_lists$probable_highest)
dataframe[dataframe$highest_category == "confirmed", 
          "species_count"] <- length(spp_lists$confirmed)

# Get species count for what's been observed
total <- length(spp_lists$all)

dataframe[dataframe$highest_category == "observed", 
          "percent_observed"] <- round(100*(length(spp_lists$observed_highest)/total))
dataframe[dataframe$highest_category == "possible", 
          "percent_observed"] <- round(100*(length(spp_lists$possible_highest)/total))
dataframe[dataframe$highest_category == "probable", 
          "percent_observed"] <- round(100*(length(spp_lists$probable_highest)/total))
dataframe[dataframe$highest_category == "confirmed", 
          "percent_observed"] <- round(100*(length(spp_lists$confirmed)/total))

# Get species count for predicted
total <- length(summer_spp)

dataframe[dataframe$highest_category == "observed", 
          "percent_predicted"] <- round(100*(length(spp_lists$observed_highest)/total))
dataframe[dataframe$highest_category == "possible", 
          "percent_predicted"] <- round(100*(length(spp_lists$possible_highest)/total))
dataframe[dataframe$highest_category == "probable", 
          "percent_predicted"] <- round(100*(length(spp_lists$probable_highest)/total))
dataframe[dataframe$highest_category == "confirmed", 
          "percent_predicted"] <- round(100*(length(spp_lists$confirmed)/total))

print(dataframe)
```



# Tests
```{r}

```


# Speed
Run the function several times to get descriptive statistics.
```{r}
# Run the function 5 times and record the runtime
time <- c()
for (i in 1:5) {
  time1 <- proc.time()
  
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```