---
title: "dev_blocks_observed_in"
author: "N.M. Tarr"
date: "2023-09-20"
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```

# Purpose
This document details a function that returns data frame of NCBA blocks within which a species was observed within or outside of a specified time period.

# Definition
```{r}
blocks_observed_in
```


# Usage

```{r}
# Get a blocks data frame with simple features
fields <- c("ID_BLOCK_CODE", "ID_EBD_NAME")
blocks_sf <- get_blocks(spatial = TRUE, fields = fields)
```

```{r}
# Species
species <- "Philadelphia Vireo"

# Pull out breeding season records
breedates <- get_breeding_dates(species, day_year = TRUE)

# Get all the observations for the species, exclude zero count records though
#fields <- c("atlas_block", "breeding_category")
observations <- get_observations(species = species, 
                        database = "AtlasCache",
                        NCBA_only = TRUE, 
                        fields = NULL) %>%
  to_EBD_format() %>%
  filter(observation_count != 0)
```

Map the blocks with any type of observation.
```{r}
pres.blocks <- blocks_observed_in(observations, start_day = 1, end_day = 365, 
                                  within = TRUE,
                                  breeding_categories = c("C4", "C3", "C2", 
                                                          "C1", ""))

pres.blocks_sf <- right_join(blocks_sf, pres.blocks, by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = pres.blocks_sf, aes(fill = common_name)) +
  ggtitle("Blocks With Observations")
```

Map the blocks with breeding season observations
```{r}
# Pull out breeding season dates
breedates <- get_breeding_dates(species, day_year = TRUE)

summer.blocks <- blocks_observed_in(observations, start_day = breedates[1], 
                                    end_day = breedates[2], 
                                    within = TRUE,
                                    breeding_categories = c("C4", "C3", "C2", 
                                                            "C1", ""))

summer.blocks_sf <- right_join(blocks_sf, summer.blocks, 
                               by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = summer.blocks_sf, aes(fill = common_name)) +
  ggtitle("Blocks With Breeding Season Observations")
```

Map blocks with nonbreeding season observations
```{r}
winter.blocks <- blocks_observed_in(observations, start_day = breedates[1], 
                                    end_day = breedates[2], 
                                    within = FALSE,
                                    breeding_categories = c("C4", "C3", "C2", 
                                                            "C1", ""))

winter.blocks_sf <- right_join(blocks_sf, winter.blocks, 
                               by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = winter.blocks_sf, aes(fill = common_name)) +
  ggtitle("Blocks With Nonbreeding Season Observations")
```

Map blocks with confirmed or probable breeding observations from any day of the year.
```{r}
confirmed <- blocks_observed_in(observations, start_day = 1, 
                                end_day = 365, 
                                within = TRUE,
                                breeding_categories = c("C3", "C4"))

confirmed_sf <- right_join(blocks_sf, confirmed, 
                   by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = confirmed_sf, aes(fill = common_name)) +
  ggtitle("Blocks with Confirmed or Probable Breeding Records")
```

Map blocks with confirmed or probable breeding observations from within breeding safe dates.
```{r}
conf.prob <- blocks_observed_in(observations, start_day = breedates[1], 
                                end_day = breedates[2], 
                                within = TRUE,
                                breeding_categories = c("C3", "C4"))

conf.prob_sf <- right_join(blocks_sf, conf.prob, 
                           by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = confirmed_sf, aes(fill = common_name)) +
  ggtitle("Blocks with Confirmed or Probable Breeding Records Within Safe Dates")
```

Map blocks with confirmed or probable breeding observations from outside of breeding safe dates.
```{r}
conf.prob <- blocks_observed_in(observations, start_day = breedates[1], 
                                end_day = breedates[2], 
                                within = FALSE,
                                breeding_categories = c("C3", "C4"))

conf.prob_sf <- right_join(blocks_sf, conf.prob, 
                           by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = confirmed_sf, aes(fill = common_name)) +
  ggtitle("Blocks with Confirmed or Probable Breeding Records Outside Safe Dates")
```

# Tests



# Speed

# Bugs
present column isn't being renamed appropriately.