---
title: "dev_block_spp_lists"
author: "N.M. Tarr"
date: "2023-10-06"
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Purpose
This document details and demonstrates a function that retrieves a list of reported species from a particular atlas block.

```{r}
block_spp_lists
```


# Usage
Get summer species lists for a block of interest (May 15 - August 15).
```{r}
block <- "35078H8SE"
spp_lists <- block_spp_lists(block = block, start_day = 135, end_day = 227,
                             within = TRUE)
```

The result of the function is a list of lists.  Retrieve the names of all the lists included.
```{r}
print(names(spp_lists))
```

Print a list of all species that have been reported.
```{r}
print(spp_lists$all)
```

Print a list of all species that have been reported as possibly breeding.
```{r}
print(spp_lists$possible)
```

Print a list of all species that have been reported as probably breeding.
```{r}
print(spp_lists$probable)
```

Print a list of all species that have been reported as confirmed breeding.
```{r}
print(spp_lists$confirmed)
```

Print a list of all species that have been reported under an "observed" breeding code or no breeding code.
```{r}
print(spp_lists$observed)
```

Print a list of all species for which observed is the highest category that has been submitted.
```{r}
print(spp_lists$observed_highest)
```

Print a list of all species for which possible is the highest category that has been submitted.
```{r}
print(spp_lists$possible_highest)
```

Print a list of all species for which probable is the highest category that has been submitted.
```{r}
print(spp_lists$probable_highest)
```

Print a list of all species for which confirmed is the highest category that has been submitted.
```{r}
print(spp_lists$confirmed)
```


# Speed
Run the function several times to get descriptive statistics.
```{r}
# Run the function 5 times and record the runtime
time <- c()
for (i in 1:5) {
  time1 <- proc.time()
  block_spp_lists(block = block, start_day = 135, end_day = 227,
                             within = TRUE)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```