---
title: "dev_describe_checklists"
author: "N.M. Tarr"
output:
  html_document:
    df_print: paged
  rmdformats::downcute: null
  word_document: default
---
```{r setup, include=TRUE, message=TRUE, warning=TRUE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```

# Purpose
This document details several functions that describes the attributes of checklists.  Each function is specified in this section and then demonstrated in the next.

## Functions
A function to make a boxplot of checklist distance in kilometers.
```{r}
effort_distance_boxplot_dev <- function(checklists){
  # Describe the distribution of effort_distance_km values as a boxplot
  # 
  # Parameters:
  # checklists -- data frame of checklists w/ effort_distance_km.
  #
  # Example:
  # effort.dist <- effort_distance_boxplot(get_all_checklists(config, 
  #                                                         drop_ncba_col=TRUE))
  # plot(effort.dist)
  boxplot <- ggplot(data=checklists) +
    geom_boxplot(mapping=aes(y=effort_distance_km, x=""), 
                 color="darkgreen", 
                 outlier.colour="blue", show.legend=TRUE) + 
    coord_flip() + 
    labs(title="",
         caption=" ") +
    ylab("Kilometers") + 
    xlab("Checklist Travel Distances") +
    scale_y_continuous(n.breaks=12)
}
```

A function to make a boxplot of checklist durations in minutes.
```{r}
duration_minutes_boxplot_dev <- function(checklists){
  # Describe the distribution of effort_minutes values as a box plot.
  # 
  # Parameters:
  # checklists -- data frame of checklists w/ effort_minutes.
  #
  # Example:
  # duration.min <- plot_checklists_coords(get_all_checklists(config, 
  #                                                        drop_ncba_col=TRUE))
  # plot(duration.min)
  boxplot <- ggplot(data=checklists) +
    geom_boxplot(mapping=aes(y=duration_minutes, x=""), 
                 color="darkblue", 
                 outlier.colour="orange", show.legend=TRUE) + 
    coord_flip() + 
    labs(title="", caption="") +
    ylab("Minutes") + 
    xlab("Checklist Durations ") +
    scale_y_continuous(n.breaks=12)
}
```

A function to make a boxplot of checklist start times.
```{r}
start_time_boxplot_dev <- function(checklists){
  # Describe the distribution of checklist start times as a box plot.
  # 
  # Parameters:
  # checklists -- data frame of checklists w/ start time_observations_started.
  #
  # Example:
  # start.box <- start_time_boxplot(get_all_checklists(config, 
  #                                                    drop_ncba_col=TRUE))
  # plot(start.box)
  # make a vector of plottable start times.
  library(hms)
  times <- checklists %>%
    filter(time_observations_started != "") %>% # empty strings cause probs.
    select(time_observations_started) %>%
    mutate(time=hour(as_hms(time_observations_started)))
  
  # make graph
  boxplot <- ggplot(data=times) + 
    geom_boxplot(mapping=aes(y=time, x=""), 
                 color="darkblue", outlier.colour="magenta", show.legend=TRUE) + 
    coord_flip() + 
    labs(title="",
         caption="") + 
    ylab("Time of Day") +
    xlab("Checklist Start Time") +
    scale_y_continuous(n.breaks=12)
}
```

A function to make a pie chart of locality types.
```{r}
# ------------------------------------------------------------------------------
locality_type_pie_dev <- function(checklists){
  # Describe the locality types present as a pie chart: whether checklists are 
  #   for hotspots, personal locations, etc.
  # 
  # Parameters:
  # checklists -- data frame of checklists w/ locality_type.
  #
  # Example:
  # locality.pie <- locality_type_pie(get_all_checklists(config, 
  #                                                        drop_ncba_col=TRUE))
  # plot(locality.pie)
  
  by_locality_type <- checklists %>%
    group_by(locality_type) %>%
    summarize(count = n())
  
  ## Print table
  #knitr::kable(by_locality_type,
  #             caption="Count of checklists per locality type")
  
  # Pie chart
  pie <- ggplot(data=by_locality_type, aes(x="", y=count, fill=locality_type)) +
    geom_bar(stat="identity", width=1) +
    coord_polar("y", start=0) +
    scale_fill_viridis_d(alpha = 1, option="D") +
    theme_void() + 
    #theme(legend.title = "Locality Type") +
    labs(title="", caption="") +
    guides(fill=guide_legend(title="Locality Type"))
}
```

A function to make a pie chart of protocol types.
```{r}
# ------------------------------------------------------------------------------
protocol_type_pie_dev <- function(checklists){
  # Describe the protocol types present as a pie chart: whether checklists are 
  #   traveling, stationary, etc.
  # 
  # Parameters:
  # checklists -- data frame of checklists w/ protocol_type.
  #
  # Example:
  # protocol.pie <- protocol_type_pie(get_all_checklists(config, 
  #                                                        drop_ncba_col=TRUE))
  # plot(protocol.pie)
  
  by_protocol_type <- checklists %>%
    group_by(protocol_type) %>%
    summarize(count = n())
  
  ## Print table
  #knitr::kable(by_protocol_type,
  #             caption="Count of checklists per protocol type")
  
  # Pie chart
  pie <- ggplot(data=by_protocol_type, aes(x="", y=count, fill=protocol_type)) +
    geom_bar(stat="identity", width=1) +
    coord_polar("y", start=0) +
    scale_fill_viridis_d(alpha = 1, option="D") +
    theme_void() + 
    labs(title="", caption="") +
    guides(fill=guide_legend(title="Protocol Type"))
}
```

A function to make a bar chart of checklist years.
```{r}
year_bar_dev <- function(checklists){
  # Summarize how many checklists were reported each year.
  # 
  # Parameters:
  # checklists -- data frame of checklists w/ year.
  #
  # Example:
  # year.bar <- year_bar(get_all_checklists(config, drop_ncba_col=TRUE))
  # plot(year.bar)
  
  barchart <- ggplot(data=checklists) +
  geom_bar(mapping=aes(x=year), 
               show.legend=FALSE) + 
  labs(title="",
       caption=" ") +
  ylab("Checklists") + 
  xlab("Year") 
}
```


# Usage
This demo requires the tidyverse packages.
```{r}
library(tidyverse)
```

Load the NCBA functions because this function relies upon the output from some of them.
```{r}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")
config <- "~/Documents/NCBA/Scripts/ncba_config.R"
```

Set the working directory to somewhere outside of the NCBA repository so that results are not saved in the repository.
```{r}
setwd("~/Documents/NCBA/species/")
```

Use the NCBA connection function to connect to the Atlas Cache (the mongodb).
```{r}
# connect to a specific collection (table)
connection <- connect_ncba_db(ncba_config = config, database = "ebd_mgmt", 
                              collection = "ebd")
```

Identify a species to investigate.
```{r}
species <- "Ovenbird"
print(species)
```

Retrieve the records for the species from the Atlas Cache
```{r}
time1 <- proc.time()

records <- get_observations(species, database = "AtlasCache", NCBA_only = TRUE,
                            EBD_fields_only = FALSE, fields = NULL, 
                            ncba_config = config)

records <- to_EBD_format(records, drop = TRUE)

# Calculate processing time
mongotime <- proc.time() - time1

# Print number of records returned
print(paste("Records returned:", nrow(records)))
print(paste("Runtime: ", mongotime[["elapsed"]]))
print(head(records))
```

Make an effort distance boxplot and display it
```{r}
distance.box <- effort_distance_boxplot_dev(records)
plot(distance.box)
```

Make an effort duration boxplot and display it
```{r}
minutes.box <- duration_minutes_boxplot_dev(records)
plot(minutes.box)
```

```{r}
start.box <- start_time_boxplot_dev(records)
plot(start.box)
```

Make a piechart showing the composition of locality types.
```{r}
locality.type <- locality_type_pie_dev(records)
plot(locality.type)
```

Make a piechart showing the composition of protocol types.
```{r}
protocol.type <- protocol_type_pie_dev(records)
plot(protocol.type)
```

Make a bar chart showing the number of checklists per year.
```{r}
year.bar <- year_bar_dev(records)
plot(year.bar)
```