---
title: "dev_get_breeding_records"
author: "N.M. Tarr"
date: "2023-12-5"
output:
  word_document: default
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Function
Retrieves the observation records with breeding/behavior codes from either the NCBA database or from a downloaded copy of the EBD.  Results can be limited to records with breeding/behavior codes of interest.  If data is requested from the Atlas Cache, then NCBA columns that are not found in the EBD databases can be dropped or retained.  Additionally, a customized list of fields can be specified to limit the columns that are included in the output data frame.
```{r}
get_breeding_records
```

# Usage
The function can target records of a certain species, project, block, or observer, as well as combinations of those four.

## Behaviors
Retrieve breeding records from the Atlas Cache with the function, but only certain fields.
```{r}
fields <- c("SAMPLING_EVENT_IDENTIFIER", "BEHAVIOR_CODE", "COMMON_NAME", 
            "ATLAS_BLOCK")

records <- get_breeding_records(behaviors = c("S7", "ON"),
                                database = "AtlasCache",
                                EBD_fields_only = TRUE,
                                fields = fields)
print(head(records, 5))
```

Retrieve breeding records from the EBD with the function.
```{r}
fields <- c("SAMPLING_EVENT_IDENTIFIER", "BEHAVIOR_CODE", "COMMON_NAME", 
            "ATLAS_BLOCK")

records <- get_breeding_records(behaviors = c("S7", "ON"),
                                database = "EBD",
                                EBD_fields_only = TRUE,
                                fields = fields)
print(head(records, 5))
```

## Block
Retrieve observations from a particular block from the Atlas Cache.
```{r}
block <- "36078C8SE"

fields <- c("ATLAS_BLOCK", "SAMPLING_EVENT_IDENTIFIER", "OBSERVER_ID", "COMMON_NAME")

timberlake <- get_breeding_records(behaviors = breeding_codes()$confirmed, 
                                   block = block, 
                                   database = "AtlasCache", 
                                   fields = fields)
print(timberlake)
```

Retrieve observations from the EBD with the function.
```{r}
timberlake <- get_breeding_records(behaviors = breeding_codes()$confirmed, 
                                   block = block, 
                                   database = "EBD", 
                                   fields = fields)
print(timberlake)
```

## Observer
Retrieve breeding records from a particular observer from the Atlas Cache.
```{r}
observer <- "obsr225263"
fields <- c("ATLAS_BLOCK", "SAMPLING_EVENT_IDENTIFIER", 
            "GLOBAL_UNIQUE_IDENTIFIER", "OBSERVER_ID", "COMMON_NAME", 
            "BEHAVIOR_CODE")
natetarr <- get_breeding_records(behaviors = breeding_codes()$confirmed, 
                                 block = NULL,
                                 observer = observer,
                                 database = "AtlasCache", 
                                 fields = fields)
print(natetarr)
```

Retrieve breeding records from the EBD with the function.
```{r}
natetarr <- get_breeding_records(behaviors = breeding_codes()$confirmed, 
                                 block = NULL,
                                 observer = observer,
                                 database = "EBD", 
                                 fields = fields)
print(natetarr)
```

## Project
Retrieve observations from a particular project from the Atlas Cache.
```{r}
fields <- c("PROJECT_CODE", "SAMPLING_EVENT_IDENTIFIER", "BREEDING_CODE",
            "GLOBAL_UNIQUE_IDENTIFIER", "COUNTY", "COMMON_NAME")

VA <- get_breeding_records(project = "EBIRD_ATL_VA", 
                           database = "AtlasCache", 
                           fields = fields)
print(head(VA, 5))
```


## All options
More than parameter can be utilized.
```{r}
fields <- c("PROJECT_CODE", "SAMPLING_EVENT_IDENTIFIER", 
            "GLOBAL_UNIQUE_IDENTIFIER", "OBSERVER_ID", "BREEDING_CODE")

obs.AC <- get_breeding_records(project = "EBIRD_ATL_NC",
                               observer = observer,
                               behaviors = c("H", "ON", "S"),
                               block = block,
                               database = "AtlasCache", 
                               fields = fields)
print(head(obs.AC, 5))
```

```{r}
obs.EBD <- get_breeding_records(project = "EBIRD_ATL_NC",
                               observer = observer,
                               behaviors = c("H", "ON", "S"),
                               block = block,
                               database = "EBD", 
                               fields = fields)
print(head(obs.EBD, 5))
```


# Tests

## From the Atlas Cache
There should only be the desired breeding codes in the data frame. If no unwanted codes are returned from this chunk, then the test is passed.
```{r}
obs <- get_breeding_records(behaviors = c("ON", "CN"))
print(unique(obs$BEHAVIOR_CODE))
```

If the project was set to NCBA, then the PROJECT_CODE field should only contain "EBIRD_ATL_NC".  This chunk returns "EBIRD_ATL_NC" if the test was passed.
```{r}
obs <- get_breeding_records(behaviors = c("CN"),
                            project = "EBIRD_ATL_NC")
print(unique(obs$PROJECT_CODE))
```

If a list of fields is provided, then the output data frame should only include the desired columns.  This chunk should return TRUE if the test is passed.
```{r}
fields <- c("SAMPLING_EVENT_IDENTIFIER", "PROJECT_CODE", "GLOBAL_UNIQUE_IDENTIFIER")

obs <- get_breeding_records(behaviors = c("CN"),
                            project = "EBIRD_ATL_NC",
                            fields = fields)
print(unique(names(obs) == fields))
```

If only fields that are found in the eBird Basic Dataset are wanted, then no NCBA derived columns should be present.  If this chunk returns FALSE, then the test is passed. 

THIS TEST FAILS because of the fields "SUBSPECIES_COMMON_NAME" and "SUBSPECIES_SCIENTIFIC_NAME".  What is their origin?
```{r}
# Get the list of NCBA columns
NCBA.columns <- nonEBD_fields() 

# Get observations
obs <- get_breeding_records(behaviors = c("CN"),
                            project = "EBIRD_ATL_NC",
                            EBD_fields_only = TRUE)

# Test whether NCBA columns are in the observation data frame
print(unique(NCBA.columns %in% names(obs)))
```

Identify any columns found in an EBD data set that are absent from the output data frame.  "checklist_id" and "age_sex" are expected to be absent because "checklist_id" is not present in observation records and "age_sex" is included in the AtlasCache as "age/sex".
```{r}
# Get an example EBD data frame
input_file <- system.file("extdata/ebd-sample.txt", package = "auk")
# output text file
output_file <- "ebd_filtered_grja.txt"
ebird_data <- input_file %>% 
  # 1. reference file
  auk_ebd() %>% 
  # 2. define filters
  auk_species(species = "Canada Jay") %>% 
  auk_country(country = "Canada") %>% 
  # 3. run filtering
  auk_filter(file = output_file, overwrite = TRUE) %>% 
  # 4. read text file into r data frame
  read_ebd()

# Get the EBD column names
EBD.columns <- names(ebird_data)

# Get a lower case version of the output data frame columns names
our.columns <- str_to_lower(names(obs))

# Is every EBD column in our column list?
print(setdiff(EBD.columns, our.columns))
```

## eBird Basic Dataset
There should only be the desired breeding codes in the data frame. If no unwanted codes are returned from this chunk, then the test is passed.
```{r}
obs <- get_breeding_records(behaviors = c("ON", "CN", "FY"),
                            database = "EBD")
print(unique(obs$behavior_code))
```

If the project was set to NCBA, then the PROJECT_CODE field should only contain "EBIRD_ATL_NC".  This chunk returns "EBIRD_ATL_NC" if the test was passed.
```{r}
obs <- get_breeding_records(behaviors = c("CN"),
                            project = "EBIRD_ATL_NC",
                            database = "EBD")
print(unique(obs$project_code))
```

If a list of fields is provided, then the output data frame should only include the desired columns.  This chunk should return TRUE if the test is passed.
```{r}
fields <- c("sampling_event_identifier", "project_code", "global_unique_identifier")

obs <- get_breeding_records(behaviors = c("CN"),
                            fields = fields,
                            database = "EBD")
print(unique(names(obs) == fields))
```


# Speed (in seconds)

## Atlas Cache
For an infrequent code
```{r}
# Run the function 3 times and record the runtime
time <- c()
for (i in 1:3) {
  time1 <- proc.time()
  get_breeding_records(behaviors = c("FS"))
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```

For a frequent code
```{r}
# Run the function 3 times and record the runtime
time <- c()
for (i in 1:3) {
  time1 <- proc.time()
  get_breeding_records(behaviors = c("H"))
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```

## EBD
For a data-poor code
```{r}
# Run the function 2 times and record the runtime
time <- c()
for (i in 1:2) {
  time1 <- proc.time()
  get_breeding_records(behaviors = c("FS"), database = "EBD")
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```

For a data-rich code
```{r}
# Run the function 2 times and record the runtime
time <- c()
for (i in 1:2) {
  time1 <- proc.time()
  get_breeding_records(behaviors = c("H"), database = "EBD")
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```