---
title: "dev_blocks_needed"
author: "N.M. Tarr"
date: "2023-10-03"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Purpose
This document defines and demonstrates a function that returns a data frame of blocks where the species was predicted to occur by GAP or eBird but the NCBA has yet to acquire observations.

# Definition
```{r}
blocks_needed
```


# Usage
Running the function returns a data frame with records for atlas blocks where the species was predicted to occur but has not yet been observed.  The data frame also has simple features so that it can be used in a map.
```{r}
species <- "Loggerhead Shrike"

# Get a data frame of blocks that need observations
needed_sf <- blocks_needed(species = "Loggerhead Shrike", 
                           source = "GAP", season = "summer",
                           database = "AtlasCache", project = "EBIRD_ATL_NC",
                           observations = NULL)
# Print 
print(head(needed_sf, 5))
```

The resulting data frame can be drawn up as a map.
```{r}
# Get a blocks data frame with simple features
fields <- c("ID_BLOCK_CODE", "ID_EBD_NAME")
blocks_sf <- get_blocks(spatial = TRUE, fields = fields)
  
# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = needed_sf, aes(fill = needed)) +
  ggtitle(str_interp("Blocks That Need ${species} Observations"))
```

The result can also be used in a map with multiple layers.  For example, it can be drawn below a layer showing the highest reported breeding category per block.  
```{r}
# Get the highest breeding category code per block
highest_sf <- highest_category(species) %>%
  drop_na()

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = needed_sf, aes(fill = needed)) + 
  geom_sf(data = highest_sf, aes(fill = highest_category)) + 
  ggtitle(str_interp("Blocks Needing Summer ${species} Observations and Highest Category"))
```

Running the get_observations function can be slow, but it can be run ahead of time and the results passed to blocks_needed in order to speed up the runtime.
```{r}
species <- "Red-breasted Nuthatch"
# Get a data frame of observations
obs <- get_observations(species = species, database = "AtlasCache",
                        project = "EBIRD_ATL_NC") %>%
      to_EBD_format()
```

Here, a data frame is passed to the function to find blocks needing winter observations.
```{r}
# Get a dataframe of blocks that need observations
needed_sf <- blocks_needed(species = species, source = "GAP", season = "winter",
                           database = database, observations = obs) 

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = needed_sf, aes(fill = needed)) + 
  ggtitle(str_interp("Blocks for Which Winter ${species} Observations are Needed"))
```

# Tests
This function relies almost entirely on other ncba_functions that were tested elsewhere, thus formal tests are not necessary.  However, various species can be explored to assure that results are as they should be.
```{r}
# Get a dataframe of blocks that need observations
needed_sf <- blocks_needed(species = "Red-breasted Nuthatch", 
                           source = "GAP", 
                           season = "summer",
                           database = database, 
                           project = "EBIRD_ATL_NC",
                           observations = obs) 

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = needed_sf, aes(fill = needed)) + 
  ggtitle("Blocks for Which Summer Red-breasted Nuthatch Observations are Needed (GAP)")
```

```{r}
# Get a dataframe of blocks that need observations
needed_sf <- blocks_needed(species = "Red-breasted Nuthatch", 
                           source = "eBird", 
                           season = "wintering",
                           database = database, 
                           project = "EBIRD_ATL_NC",
                           observations = obs) 

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = needed_sf, aes(fill = needed)) + 
  ggtitle("Blocks for Which Red-breasted Nuthatch Winter Observations are Needed (eBird)")
```

```{r}
# Get a data frame of blocks that need observations
needed_sf <- blocks_needed(species = "Bald Eagle", 
                           source = "GAP", 
                           season = "winter",
                           database = database, 
                           project = NULL,
                           observations = obs) 

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = needed_sf, aes(fill = needed)) + 
  ggtitle("Blocks for Which Bald Eagle Winter Observations are Needed (GAP)")
```

```{r}
# Get a dataframe of blocks that need observations
needed_sf <- blocks_needed(species = "Bald Eagle", 
                           source = "eBird", 
                           season = "breeding",
                           database = database, 
                           project = "EBIRD_ATL_NC",
                           observations = obs) 

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = needed_sf, aes(fill = needed)) + 
  ggtitle("Blocks for Which Bald Eagle Breeding Observations are Needed (eBird)")
```


# Speed
Run the function 5 times and record the runtime
```{r}
time <- c()
for (i in 1:5) {
  time1 <- proc.time()
  blocks_needed(species = species, source = "GAP", season = "winter",
                           database = database, project = "EBIRD_ATL_NC",
                           observations = obs)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```