---
title: "dev_get_checklists"
author: "N.M. Tarr"
date: "2023-08-28"
output: html_document
---

```{r setup, include=TRUE, message=TRUE, warning=TRUE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Function
```{r}
get_checklists <- function(database = "AtlasCache", EBD_fields_only = TRUE,
                           NCBA_only = TRUE, fields = NULL, 
                           observer = NULL, block = NULL, EBD_format = FALSE){
  # Get a data frame of checklists from the AtlasCache.  Use get_observations()
  #   instead if you want species observation records.
  # 
  # Description:
  # This function enables the retrieval of checklist records from the AtlasCache
  #   or the EBD.  It offers options to restrict records to those with a 
  #   particular project id, block id, or observer id.  Additionally, one can 
  #   specify a subset of fields to return or drop non-EBD fields.  to apply the
  #   to_EBD_format() function to the result of an AtlasCache query, set the
  #   EBD_format variable to TRUE.
  #
  #
  # Parameters:
  # database -- either "EBD" for a downloaded eBird database or "AtlasCache" for
  #   the NCBA mongodb.
  # EBD_fields_only -- whether to include non-EBD, Atlas Cache fields in the output 
  #   data frame. TRUE or FALSE and defaults to FALSE. This argument is set to 
  #   FALSE if the fields argument in not NULL. When database is set to EBD,
  #   this parameter is obsolete.
  # NCBA_only -- whether to exclude non-NCBA project records.  This argument 
  #   is set to FALSE if the fields argument in not NULL.
  # fields -- a list of fields to return, excluding those not listed.  This 
  #   parameter offers no speed benefit with EBD sampling datasets.
  # observer -- an observer id to filter based on. For example, "obs421398"
  # block -- a block id to filter based on. For example, "34080G3NW"
  #
  # Notes:
  # - Data frame output when setting database to "AtlasCache" may require 
  #     additional wrangling with the to_EBD_format function before subsequent
  #     functions can be used.
  
  library(tidyverse)
  library(auk)
  
  # Set the working directory
  if (is.null(work_dir) == FALSE) {
    setwd(work_dir)
  }
  
  # If a list of fields is provided, set EBD_fields_only to FALSE
  if (is.null(fields) == FALSE) {
    EBD_fields_only = FALSE
  }
  
  if (database == "AtlasCache") {
    # Connect to the NCBA database
    connection <- connect_ncba_db("ebd_mgmt", "ebd")
    
    # Define a query sequentially.  First, address NCBA_only.
    if (NCBA_only == FALSE) {
      query <- '{}'
    } else {
      query <- '{"PROJECT_CODE" : "EBIRD_ATL_NC"}'
    }
    
    # Next, address observer
    if (is.null(observer) == FALSE) {
      new_end <- str_interp(', "OBSERVER_ID" : "${observer}"}')
      query <- paste0(substr(query, 1, nchar(query)-1), new_end)
    } else {
      query <- query
    }
    
    # Next, address block id
    if (is.null(block) == FALSE) {
      new_end <- str_interp(', "ATLAS_BLOCK" : "${block}"}')
      query <- paste0(substr(query, 1, nchar(query)-1), new_end)
    } else {
      query <- query
    }
    
    # Define a filter that excludes the observation column...
    if (is.null(fields) == TRUE) {
      fields <- '{"OBSERVATIONS" : false}'
      
      # Identify AC fields for omission
      AC.fields <- nonEBD_fields()
      
      # Build a field string for the query if necessary
      if (EBD_fields_only == TRUE) {
        # Convert the list of field names to a mongolite filter string
        fields_string <- paste0('{', paste0('"', AC.fields, '" : false', 
                                            collapse = ', '))
        
        # Redefine fields so that it can be pasted with the filter string
        fields <- ', "OBSERVATIONS" : false}'
        
        # Combine with the existing fields string
        fields <- paste0(fields_string, fields)
      }
      
      # ... but if fields are provided, use those as a filter
    } else {
      # Convert the list of field names to a mongolite filter string
      fields_string <- paste0('{', paste0('"', fields, '" : true', 
                                          collapse = ', '), '}')
      fields <- c(fields_string)
    }
    
    # Retrieve the checklists
    checklists <- connection$find(query = query, fields = fields)
    
    # Reformat to EBD if requested
    if (EBD_format == TRUE) {
      checklists <- checklists %>% to_EBD_format()
    }
  }
  
  
  if (database == "EBD") {
    library(auk)
    
    # Condition next action on whether NCBA records only are desired.
    if (NCBA_only == TRUE) {
      # Read in sampling data frame with auk
      sampling <- EBD_sampling %>%
        auk_sampling() %>%
        auk_project("EBIRD_ATL_NC") %>%
        auk_filter(file = "TMP_EBD.txt", overwrite = TRUE) %>%
        read_sampling() %>%
        data.frame()
    } else {
      sampling <- EBD_sampling %>%
        auk_sampling() %>%
        auk_filter(file = "TMP_EBD.txt", overwrite = TRUE) %>%
        read_sampling() %>%
        data.frame()
    }
    
    # Pull out desired block and observer
    if (is.null(observer) == FALSE) {
      sampling <- filter(sampling, observer_id == observer)
    } else {
      sampling <- sampling
    }
    
    if (is.null(block) == FALSE) {
      sampling <- filter(sampling, atlas_block == block)
    } else {
      sampling <- sampling
    }

    # Subset the columns
    if (is.null(fields) == FALSE) {
      checklists <- select(sampling, fields)
    } else {
      checklists <- sampling
    }
  }
  
  return(checklists)
}
```


# Usage
Retrieve all checklists from the Atlas Cache.
```{r}
checklists <- get_checklists(database = "AtlasCache")
print(head(checklists))
```

# Tests

The default data frame returned from the AtlasCache should not include the OBSERVATIONS field.  This chunk should return FALSE to pass the test.
```{r}
checklists <- get_checklists(database = "AtlasCache")

print("OBSERVATIONS" %in% names(checklists))
```

Records from the Atlas Cache should also not include the nonEBD fields when EBD_fields_only is set to TRUE, which is the default.  Thus, the only value returned by this chunk should be FALSE to pass this test.
```{r}
print(unique(nonEBD_fields() %in% names(checklists)))
```

On the other hand, if EBD_fields_only is set to FALSE, then non-EBD fields should be present in a returned data frame and the test from above should return TRUE in this chunk.
```{r}
checklists <- get_checklists(database = "AtlasCache",
                                 EBD_fields_only = FALSE)

print(unique(nonEBD_fields() %in% names(checklists)))
```

There also shouldn't be an OBSERVATIONS column in the data frame.  This chunk should return FALSE to pass the test
```{r}
print("OBSERVATIONS" %in% names(checklists))
```


Utilizing the fields parameters should limit the fields returned.  Thus, the column names from the output data frame should be the same as the fields argument and this chunk should return TRUE to pass the test.
```{r}
# Make a list of fields for a test.  Choose some from EBD and others from NCBA-specific fields.
testfields <- c("_id", "LOCALITY_ID", "SAMPLING_EVENT_IDENTIFIER", 
                "NCBA_SEASON", "GEOM") 

checklists2 <- get_checklists(database = "AtlasCache", fields = testfields)

unique(sort(names(checklists2)) == sort(testfields))
```

EBird checklists that were submitted with the NCBA portal are assigned the value "EBIRD_ATL_NC" in the PROJECT_CODE field.  Therefore, using the NCBA_only parameter should return a data frame with no values other than EBIRD_ATL_NC in the PROJECT_CODE field.  This test is passed if the chunk returns "EBIRD_ATL_NC".
```{r}
checklists3 <- get_checklists(database = "AtlasCache", NCBA_only = TRUE)
print(unique(checklists3$PROJECT_CODE))
```

Perform the same test, but with the EBD database.  WARNING: this is very slow.
```{r}
# checklists4 <- get_checklists(database = "EBD", NCBA_only = TRUE)
# print(unique(checklists4$project_code))
```


# Speed (in seconds)
Describe runtimes for queries of the Atlas Cache with ncba_only = TRUE
```{r}
# Run the function 5 times and record the runtime
time <- c()
for (i in 1:5) {
  time1 <- proc.time()
  get_checklists(database = "AtlasCache", NCBA_only = TRUE)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```


Measure the runtime of an EBD query, which tend to be very slow, so only do that one
once.
```{r}
# time1 <- proc.time()
# get_checklists(database = "EBD", NCBA_only = TRUE)
# print(proc.time() - time1)
```




