---
title: "dev_describe_records"
author: "N.M. Tarr"
output:
  word_document: default
  rmdformats::downcute: null
  html_document:
    df_print: paged
---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```

# Purpose
This document details several functions that describes the attributes of checklists.  Each function is specified in this section and then demonstrated in the next.

## Functions
A function to make a boxplot of checklist distance in kilometers.
```{r}
effort_distance_boxplot
```

A function to make a boxplot of checklist durations in minutes.
```{r}
duration_minutes_boxplot
```

A function to make a boxplot of checklist start times.
```{r}
start_time_boxplot
```

A function to make a pie chart of locality types.
```{r}
locality_type_pie
```

A function to make a pie chart of protocol types.
```{r}
# ------------------------------------------------------------------------------
protocol_type_pie
```

A function to make a bar chart of checklist years.
```{r}
year_bar
```


# Usage
This demo requires the tidyverse packages.
```{r}
#library(tidyverse)
```


Identify a species to investigate.
```{r}
species <- "Ovenbird"
print(species)
```

Retrieve the records for the species from the Atlas Cache
```{r}
records <- get_observations(species = species, database = "AtlasCache", 
                            project = "EBIRD_ATL_NC", EBD_fields_only = FALSE, 
                            fields = NULL) %>%
  to_EBD_format(drop = TRUE)
```

Make an effort distance boxplot and display it
```{r}
distance.box <- effort_distance_boxplot(records)
plot(distance.box)
```

Make an effort duration boxplot and display it
```{r}
minutes.box <- duration_minutes_boxplot(records)
plot(minutes.box)
```

```{r}
start.box <- start_time_boxplot(records)
plot(start.box)
```

Make a piechart showing the composition of locality types.
```{r}
locality.type <- locality_type_pie(records)
plot(locality.type)
```

Make a piechart showing the composition of protocol types.
```{r}
protocol.type <- protocol_type_pie(records)
plot(protocol.type)
```

Make a bar chart showing the number of checklists per year.
```{r}
year.bar <- year_bar(records)
plot(year.bar)
```