---
title: "Breeding Boxplot Function"
author: "N.M. Tarr"
output: 
  html_document:
    df_print: paged
    code_folding: hide
  rmdformats::downcute:
---
```{r setup, include=FALSE, message=TRUE, warning=TRUE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Purpose
This document details a function that generates a boxplot of breeding codes with some customization options and supports the ability to view the checklist behind individual data points by clicking on them in the figure.  The function offers several options as parameters.

species -- common name of the species
data -- data frame of ebird or NCBA data
type -- whether to create an interactive plot that supports opening
   checklist URLs by clicking on data points in the figure, a non-interactive plot, or a non-interactive plot separated by ecoregion.
pallet -- specify an RColorBrewer pallet (multiple colors), or a single
   color (name or hex) for the figure.
omit_codes -- specify evidence codes not be plotted.
lump -- an option to lump breeding codes into fewer categories.
drop -- TRUE or FALSE whether to include unreported codes in the plot

# Function Definition
```{r}
breeding_boxplot
```


# Usage
Identify a species to investigate.
```{r}
species <- "Red-breasted Nuthatch"
print(species)
```

Retrieve the records for the species from the Atlas Cache
```{r}
records <- get_observations(species, 
                            NCBA_only = TRUE, 
                            EBD_fields_only = FALSE) %>% 
  to_EBD_format(drop=FALSE)
```

Make a non-interactive boxplot without lumping codes together
```{r}
breeding_boxplot(species, records, type = "non-interactive")
```

Make an interactive boxplot without lumping codes together
```{r}
breeding_boxplot(species, records)
```

Create an interactive boxplot with some of the codes lumped together.
```{r}
lump <- list(S = c("S", "S7", "M"), O = c("NULL", "F", "O", "NC"))

breeding_boxplot(species, records, lump=lump)
```

Create an interactive boxplot that omits some codes and uses a different colormap.
```{r}
omit <- c("S7", "M", "NC", "H", "T", "P", "C", "CN", "N", "A", "NB", 
          "FS")
breeding_boxplot(species, records, type = "interactive", pallet="Set3",
                     omit_codes=omit, lump=NULL, drop=TRUE)
```

Make a figure that has a boxplot for each ecoregion.
```{r}
breeding_boxplot(species = species, data = records, type = "ecoregional",
                     lump = breeding_codes(lumped = TRUE))
```


# Tests
Fully programmed tests are not feasible because the function returns a figure, but we can summarize the aspects of the dataset behind the figure and visually compare them to what the figure shows.  From the cells above, the input data frame is named "records".

1. Lump all codes into two categories as an extreme example to make it clear that they work. Use category names that aren't breeding codes in order to show they can be handled.
```{r}
lump <- list(X = c("S", "S7", "M", "H", "NULL", "P", "F", "UN", "PE", "NY"), 
             Z = c("A", "N", "CN", "T", "NB","FS", "FL", "CF", "FY", "C",
                   "ON", "NE", "B"))

breeding_boxplot(species, records, type = "non-interactive", lump=lump)

breeding_boxplot(species, records, type = "interactive", lump=lump)

breeding_boxplot(species, records, type = "ecoregional", lump=lump)


```

A standard lumping list is available for the primary categories.
```{r}
lumped_codes <- breeding_codes()
breeding_boxplot(species, records, type = "interactive", lump=lumped_codes)
```

2. Omit all but three codes as an extreme example to make sure things work.  
```{r}
omit <- c("S", "M", "H", "NULL", "P", "A", "N", "CN", "T", "NB","FS", "CF", "C")

breeding_boxplot(species, records, type = "non-interactive", omit_codes = omit)

breeding_boxplot(species, records, type = "interactive", omit_codes = omit)

breeding_boxplot(species, records, type = "ecoregional", omit_codes = omit)

```

3. Check that the pallet argument works by assigning a non-default value.
```{r}
breeding_boxplot(species, records, pallet="Spectral")
breeding_boxplot(species, records, pallet="Paired")

breeding_boxplot(species, records, type="non-interactive", pallet="Spectral")
breeding_boxplot(species, records, type="non-interactive", pallet="Paired")
```

4. Check that all breeding codes that are present in the figure when drop is set to "FALSE".
```{r}
breeding_boxplot(species, records, type="non-interactive", drop=FALSE)
breeding_boxplot(species, records, drop=FALSE)
breeding_boxplot(species, records, type="ecoregional", drop=FALSE)
```

5. Check that all breeding codes that are present in records are included in the y-axis, when drop is set to "TRUE".  Note that "" is replaced with "NULL" by the function.
```{r}
# Print the unique breeding codes in records and the figure
print(unique(records$breeding_code))
breeding_boxplot(species, records, type="interactive", drop=TRUE)
breeding_boxplot(species, records, type="non-interactive", drop=TRUE)
breeding_boxplot(species, records, type="ecoregional", drop=TRUE)

```

# Speed Test
Run the function several times to get descriptive statistics.
```{r}
# Run the function 10 times and record the runtime
time <- c()
for (i in 1:10) {
  time1 <- proc.time()
  breeding_boxplot(species, records, type = "non-interactive")
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```

Results of the speed test were...

 Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
0.0220  0.0240  0.0250  0.0261  0.0260  0.0470 