---
title: "dev_protocol_table"
author: "N.M. Tarr"
date: "2023-10-06"
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Purpose
This document details a functio to summarize the protocol types of a data frame of records.

# Function
```{r}
protocol_table
```

# Usage
Get a data frame of checklist records.
```{r}
records <- get_checklists(observer = "obsr1000095", project = NULL) %>%
  to_EBD_format()
```

Run the function with the records.
```{r}
protocol_table(records)
```

Summarise the protocol type of observation records.
```{r}
willow <- get_observations(species = "Willow Flycatcher")  %>%
  to_EBD_format()

protocol_table(willow)
```


# Tests
Make sure all the protocol types are represented.  This block should return TRUE to pass the test.
```{r}
output <- unique(protocol_table(willow)$protocol_type)

input <- unique(willow$protocol_type)

print(unique(union(output, input) == output))
```


# Speed
Run the function several times to get descriptive statistics.
```{r}
# Get the data
records <- get_checklists(observer = "obsr1000095", project = NULL) %>%
to_EBD_format()

# Run the function 10 times and record the runtime
time <- c()
for (i in 1:10) {
  time1 <- proc.time()
  protocol_table(records)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```