---
title: "Breeding Boxplot Function"
author: "N.M. Tarr"
output: 
  html_document:
    df_print: paged
    code_folding: hide
  rmdformats::downcute:
---
```{r setup, include=TRUE, message=TRUE, warning=TRUE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```

# Purpose
This document details a function that generates a boxplot of breeding codes with some customization options and supports the ability to view the checklist behind individual data points by clicking on them in the figure.  The function offers several options as arguments.

interactive -- whether to create an interactive plot that supports opening
   checklist URLs by clicking on data points in the figure.
pallet -- specify an RColorBrewer pallet (multiple colors), or a single
   color (name or hex) for the figure.
omit_codes -- specify evidence codes not be plotted.
lump -- an option to lump breeding codes into fewer categories.
drop -- TRUE or FALSE whether to include unreported codes in the plot

# Function Definition
```{r}
# ------------------------------------------------------------------------------
breeding_boxplot_dev <- function(species, data, interactive=TRUE, 
                                 pallet="Paired", omit_codes=NULL,
                                 lump=NULL, drop=TRUE, cex.x.axis = 0.9, 
                                 cex.y.axis = 0.8) {
  # Produces a boxplot of breeding codes over calendar day.
  #
  # Description:
  #   Produces a boxplot of breeding codes with some customization options.  
  #     Copied from the wbbii_tools repo and altered.
  # 
  # Arguments:
  # species -- common name of the species
  # data -- data frame of ebird or NCBA data
  # interactive -- whether to create an interactive plot that supports opening
  #   checklist URLs by clicking on data points in the figure.
  # pallet -- choose a named RColorBrewer pallet (multiple colors), or a single
  #   color (name or hex); see brewer.pal.info for list and 
  #   display.brewer.all() to view all pallets
  # omit_codes -- a vector of evidence codes not be plotted. For example, 
  #   c("PE", "UN")
  # lump -- a list of named vectors where the vector name is used to place all
  #   codes in the corresponding vector (e.g. 'S = c("S", "S7", "M")' replaces
  #   all "S", "S7", and "M" with "S"). Note that any code that is not already in
  #   variable "codelevels" in function "chronplot" (below) will need to be added
  #   there.
  # drop -- TRUE or FALSE whether to include unreported codes in the plot
  library(lubridate)
  library(grid)
  library(gridBase)
  library(RColorBrewer)
  library(ggiraph)
  library(ggplot2)
  
  # Data prep ------------------------------------------------------------------
  ebird <- data # This should eventually be removed and ebird renamed.
  
  # replace breeding code entries "" with NULL
  ebird["breeding_code"][ebird["breeding_code"] == ""] <- "NULL"

  # put all dates within the same year -- ignores leap year
  ebird$observation_date <- sub("^20\\d\\d", "2050", ebird$observation_date)
  
  # remove white space from evidence codes
  ebird$breeding_code <- trimws(ebird$breeding_code)
  
  # make obsdate a date object
  ebird$obsdate <- as.Date(ebird$observation_date, "%Y-%m-%d")
  
  # Manage Breeding Codes ------------------------------------------------------
  # set drop to true if lump is used
  if (is.null(lump) == FALSE) {
    drop <- TRUE
  }
  
  # set drop to true if no plot codes is used
  if (is.null(omit_codes) == FALSE) {
    drop <- TRUE
  }
  
  # specificy breeding codes and preferred plotting order
  # this vector will need updating if any new codes are introduced via "lump".
  codelevels <- c("H", "S", "S7", "M", "T", "P", "C", "B", "CN", "NB", "A", "N",
                  "DD", "ON", "NE", "FS", "CF", "NY", "FY", "FL", "PE", "UN",
                  "F", "O", "NC", "NULL")
  # http://stackoverflow.com/questions/19681586/ordering-bars-in-barplot

  # add any new codes from the lump categories    
  if (is.null("lump") == FALSE) {
      codelevels <- c(codelevels, names(lump)[! names(lump) %in% codelevels])
  }

  # warn of unknown breeding codes in the data
  if (! all(ebird$breeding_code %in% codelevels)) {
    warn <- paste("Not all eBird codes (breeding_code) for",
                  species, "are in codelevels")
    warning(warn)
  }

  # add in unreported breeding codes to the data if drop is set to FALSE
  if (drop == FALSE) {
    # add rows with breeding codes from codelevels that are not present in ebird, 
    # but are present in codelevels.  Leave all field blank except for breeding_code.
    # get missing codes
    missing_codes <- codelevels[! codelevels %in% ebird$breeding_code]

    # make a dataframe with same columns as ebird where all fields are blank except
    # for breeding_code
    missing <- data.frame(matrix(ncol = ncol(ebird), 
                                 nrow = length(missing_codes)))
    names(missing) <- names(ebird)
  
    # add missing_codes to breeding_codes
    missing$breeding_code <- missing_codes

    # add missing codes to ebird
    ebird <- rbind(ebird, missing)
    
    # make breeding codes factors so they are ordered correctly
    ebird <- ebird %>% 
    mutate(breeding_code = factor(ebird$breeding_code, 
                                  levels = codelevels, ordered = TRUE))
  }
  
  # if drop is set to TRUE, use present breeding codes as the code levels
  #   but maintain the desired order.
  if (drop == TRUE) {
    # remove unwanted evidence codes
    if (is.null("omit_codes") == FALSE) {
      ebird <- ebird[! ebird$breeding_code %in% omit_codes, ]
    }
    
    # lump evidence codes if lump has been set
    for (i in seq_along(lump)) {
        indx <- ebird$breeding_code %in% lump[[i]]
        ebird[indx, "breeding_code"] <- names(lump)[i]
    }
    
    # make breeding codes factors so they are ordered correctly
    codelevels <- codelevels[codelevels %in% ebird$breeding_code]
    
    if (is.null("omit_codes") == FALSE) {
      codelevels <- codelevels[! codelevels %in% omit_codes]
    }

    ebird <- ebird %>% 
    mutate(breeding_code = factor(ebird$breeding_code, levels = codelevels, 
                                  ordered = TRUE))
  }
  
  
  # Colors ---------------------------------------------------------------------
  # associate colors with codelevels
  if (pallet %in% rownames(brewer.pal.info)) {
    n <- brewer.pal.info[pallet, "maxcolors"]
    codecolors <- colorRampPalette(brewer.pal(n, pallet))(length(codelevels))
  } else {
    codecolors <- rep(pallet, length(codelevels))
  }

  # colors 
  names(codecolors) <- codelevels
  
  # add column for color
  ebird$col <- codecolors[ebird$breeding_code]

  # Non-interactive plot -------------------------------------------------------
  if (interactive == FALSE) {
    # plot "empty" box plot
    boxplot(obsdate ~ breeding_code, horizontal = TRUE, 
            cex.axis = cex.y.axis, xaxt = "n", data = ebird, border = "white", 
            main = species, las = 2, xlab = "Calendar Day", 
            ylab = "Breeding Code", show.names = TRUE,
            na.action = na.pass)
    
    have_dates <- subset(ebird, is.na(observation_date) == FALSE)
    date0 <- round_date(min(have_dates$obsdate), "month")
    date1 <- round_date(max(have_dates$obsdate), "month")
    labels <- seq(from = date0, to = date1, by = "month")

    if (length(unique(month(have_dates$obsdate))) == 1) {
      labels <- c(min(have_dates$obsdate), max(have_dates$obsdate))
      labels <- unique(labels)  # in case there's only one obs
    } else {
      # limit labels to those within observed range
      ##int <- interval(min(have_dates$obsdate), max(have_dates$obsdate))
      ##labels <- labels[labels %within% int]

      if (nrow(have_dates) > 1 && length(labels) == 1) {
        labels <- unique(c(min(have_dates$obsdate), max(have_dates$obsdate)))
      }
    }

    # use format "%m/%d" for e.g. 06/01
    # use format "%b %d" for e.g. "Aug 23"
    names(labels) <- format(labels, "%b %d")

    vps <- baseViewports()
    pushViewport(vps$inner, vps$figure, vps$plot)

    # label x axis; set font size in gpar(cex = relative_fontsize);
    # grid.text is can be hard to follow but allows for arbitrary rotation of
    # x labels
    grid.text(names(labels), x = unit(labels, "native"),
              y = unit(-0.7, "lines"), just = "right", rot = 65,
              gp = gpar(cex = cex.x.axis))
    popViewport(3)

    # add tick marks
    axis(1, labels, labels = FALSE)

    # uncomment this to label the x axis a second time for sanity check
    # because grid.text can be difficult to understand
    # axis(1, labels, format(labels, "%m/%d"), col.axis = "red", las = 2)

   # select colors for stripchart
   # should be able to use "codecolors[levels(ebird$breeding_code)]",  but
   # that's giving an issue matching the empty string...
   #col <- codecolors[names(codecolors) %in% levels(ebird$breeding_code)]
   col <- unique(ebird$col)

   stripchart(obsdate ~ breeding_code, data = ebird, vertical = FALSE,
              method = "jitter", pch = 16, col = col, add = TRUE, 
              na.action = na.pass)

   # plot
   boxplot(obsdate ~ breeding_code, horizontal = TRUE,  col = "#F5F5F500",
           yaxt = "n", xaxt = "n", data = ebird, add = TRUE, 
           na.action = na.pass)
  }
  
  # Interactive plot -----------------------------------------------------------
  if (interactive == TRUE) {
    ebird$front <- 'https://ebird.org/checklist/'
    ebird$ChecklistLink <- with(ebird, paste0(front, sampling_event_identifier))

    # ggiraph code for boxplot and interactive points
    gg_point = ggplot(data = ebird) +
      labs(y="Breeding Code", x="Calendar Day") +
      geom_boxplot(aes(x = obsdate, y = breeding_code)) +
      geom_point_interactive(aes(x = obsdate, y = breeding_code, color = col, 
                                 tooltip = obsdate, data_id = obsdate,
                                 onclick=paste0('window.open("', ChecklistLink,
                                               '", "_blank")')),
                                 show.legend = FALSE, 
                                 position = position_jitter(width = .2, 
                                                            height = .2)) +
                                 theme_minimal() + labs(title = species)
  
    girafe(ggobj = gg_point, width_svg=10, 
           options = list(opts_sizing(rescale = TRUE)))
  }
}
```

# Usage
This demo requires the tidyverse packages.
```{r}
library(tidyverse)
```

Load the NCBA functions because this function relies upon the output from some of them.
```{r}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")
config <- "~/Documents/NCBA/Scripts/ncba_config.R"
```

Set the working directory to somewhere outside of the NCBA repository so that results are not saved in the repository.
```{r}
setwd("~/Documents/NCBA/species/")
```

Use the NCBA connection function to connect to the Atlas Cache (the mongodb).
```{r}
# connect to a specific collection (table)
connection <- connect_ncba_db(ncba_config = config, database = "ebd_mgmt", 
                              collection = "ebd")
```

Identify a species to investigate.
```{r}
species <- "Willow Flycatcher"
print(species)
```

Retrieve the records for the species from the Atlas Cache
```{r}
time1 <- proc.time()

# execute a query
query <- str_interp('{"OBSERVATIONS.COMMON_NAME":"${species}"}')

nc_data <- connection$find(query) %>%
  unnest(cols = (c(OBSERVATIONS))) %>% # Expand observations
  filter(COMMON_NAME == species)

# format columns to the standard analysis format (ebd format)
records <- to_ebd_format(nc_data, drop=FALSE)

# Calculate processing time
mongotime <- proc.time() - time1

# Print number of records returned
print(paste("Records returned:", nrow(records)))
print(paste("Runtime: ", mongotime[["elapsed"]]))
```

Make a non-interactive boxplot without lumping codes together
```{r}
breeding_boxplot_dev(species, records, interactive=FALSE)
```

Make an interactive boxplot without lumping codes together
```{r}
breeding_boxplot_dev(species, records)
```

Create an interactive boxplot with some of the codes lumped together.
```{r}
lump <- list(S = c("S", "S7", "M"), O = c("NULL", "F", "O", "NC"))

breeding_boxplot_dev(species, records, lump=lump)
```

Create an interactive boxplot that omits some codes and uses a different colormap.
```{r}
omit <- c("S7", "M", "NC", "H", "T", "P", "C", "CN", "N", "A", "NB", 
          "FS")
breeding_boxplot_dev(species, records, interactive=TRUE, pallet="Set3",
                 omit_codes=omit, lump=NULL, drop=TRUE)
```

# Tests
Fully programmed tests are not feasible because the function returns a figure, but we can summarize the aspects of the dataset behind the figure and visually compare them to what the figure shows.  From the cells above, the input data frame is named "records".

1. Lump all codes into two categories as an extreme example to make it clear that they work. Use category names that aren't breeding codes in order to show they can be handled.
```{r}
lump <- list(X = c("S", "S7", "M", "H", "NULL", "P"), Z = c("A", "N", "CN", "T", 
                                                        "NB","FS", "FL", "CF",
                                                        "FY", "C"))

breeding_boxplot_dev(species, records, interactive=FALSE, lump=lump)

breeding_boxplot_dev(species, records, interactive=TRUE, lump=lump)

```
2. Omit all but three codes as an extreme example to make sure things work.  
```{r}
omit <- c("S", "M", "H", "NULL", "P", "A", "N", "CN", "T", "NB","FS", "CF", "C")

breeding_boxplot_dev(species, records, interactive=FALSE, omit_codes = omit)

breeding_boxplot_dev(species, records, interactive=TRUE, omit_codes = omit)
```

3. Check that the pallet argument works by assigning a non-default value.
```{r}
breeding_boxplot_dev(species, records, interactive=TRUE, pallet="Spectral")
breeding_boxplot_dev(species, records, interactive=TRUE, pallet="Paired")

breeding_boxplot_dev(species, records, interactive=FALSE, pallet="Spectral")
breeding_boxplot_dev(species, records, interactive=FALSE, pallet="Paired")
```

4. Check that all breeding codes that are present in the figure when drop is set to "FALSE".
```{r}
breeding_boxplot_dev(species, records, interactive=FALSE, drop=FALSE)
breeding_boxplot_dev(species, records, interactive=TRUE, drop=FALSE)
```

5. Check that all breeding codes that are present in records are included in the y-axis, when drop is set to "TRUE".  Note that "" is replaced with "NULL" by the function.
```{r}
# Print the unique breeding codes in records and the figure
print(unique(records$breeding_code))
breeding_boxplot_dev(species, records, interactive=FALSE, drop=TRUE)
breeding_boxplot_dev(species, records, interactive=TRUE, drop=TRUE)
```
