---
title: "dev_species_block_map"
author: "N.M. Tarr"
date: "2023-09-12"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```

# IN PROGRESS

# Purpose
This document defines and demonstrates a function that supports the following data contract task:

"Map of blocks where the species has been predicted to occur, but has not yet been reported, including highest coded observation per species in block and distinguished by breeding/non-breeding season based on best available safe dates, GAP predictions, or other appropriate metrics."

# Definition
```{r}
blocks_observed_in <- function(what_weeks, breeding_categories) 
  {
  # Returns a dataframe of blocks where the species was observed.
  #
  # Description:
  # 
  #
  # Parameters:
  # species -- the common name of the species of interest.
  # source -- who's prediction you want: "GAP" or "eBird"
  # which_weeks -- 
  # breeding_categories -- 
  df <- NULL

  
  return(df)
}
```


```{r}
season <- "summer" # SUMMER, WINTER or BREEDING, WINTERING
species <- "Red-breasted Nuthatch"
source <- "GAP" # "eBird"
```

```{r}
# Get a blocks data frame with simple features
fields <- c("ID_BLOCK_CODE", "ID_EBD_NAME")
blocks_sf <- get_blocks(spatial = TRUE, fields = fields)
```

```{r}
# Get all the observations for the species
fields <- c("atlas_block", "observation_count")
obs <- get_observations(species = species, database = "AtlasCache",
                        NCBA_only = TRUE, fields = NULL) %>%
  to_EBD_format() %>%
  filter(observation_count != 0)

# Pull out the blocks with observations
pres.blocks <- obs %>%
  select(fields) %>%
  distinct()
```


```{r}
pres.blocks_sf <- right_join(blocks_sf, pres.blocks, by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = pres.blocks_sf, aes(fill = "blue")) +
  ggtitle("Blocks With Observations")
```

```{r}
# Pull out breeding season records
breedates <- get_breeding_dates(species, day_year = TRUE)
```

```{r}
nonbreeding.obs <- obs %>%
  filter(yday(observation_date) < breedates[[1]] | yday(observation_date) > breedates[[2]])

nonbreeding_sf <- right_join(blocks_sf, nonbreeding.obs, 
                   by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = GAP.winter, aes(fill = GAP_SPP_WINTER)) +
  geom_sf(data = nonbreeding_sf, aes(fill = common_name)) +
  ggtitle("GAP Winter Range and Winter NCBA Observations")
```


```{r}
breeding.obs <- obs %>%
  filter(yday(observation_date) > breedates[[1]] & yday(observation_date) < breedates[[2]])

breeding_sf <- right_join(blocks_sf, breeding.obs, 
                   by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = breeding_sf, aes(fill = common_name)) +
  ggtitle("Breeding Season Observations")
```


```{r}
# Confirmed
confirmed <- obs %>%
  filter(yday(observation_date) > breedates[[1]] & yday(observation_date) < breedates[[2]]) %>%
  filter(breeding_category == "C4")

confirmed_sf <- right_join(blocks_sf, confirmed, 
                   by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = confirmed_sf, aes(fill = common_name)) +
  ggtitle("Blocks with Confirmed Breeding")
```

```{r}
# Probable
probable <- obs %>%
  filter(yday(observation_date) > breedates[[1]] & yday(observation_date) < breedates[[2]]) %>%
  filter(breeding_category == "C3")

probable_sf <- right_join(blocks_sf, probable, 
                   by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = probable_sf, aes(fill = common_name)) +
  ggtitle("Blocks with Probable Breeding")
```

```{r}
# Possible
possible <- obs %>%
  filter(yday(observation_date) > breedates[[1]] & yday(observation_date) < breedates[[2]]) %>%
  filter(breeding_category == "C2")

possible_sf <- right_join(blocks_sf, possible,
                   by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = possible_sf, aes(fill = common_name)) +
  ggtitle("Blocks with Possible Breeding")
```


```{r}
# FIND THE HIGHEST CODE PER BLOCK
highest_sf <- highest_category(species, dataframe = obs)

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = GAP.summer, aes(fill = present)) + 
  geom_sf(data = highest_sf, aes(fill = highest_category)) + 
  ggtitle("Highest Reported Breeding Category and GAP Summer")
```





# Usage
Map the blocks with any type of observation.
```{r}
pres.blocks <- blocks_observed_in(observations, start_day = 1, end_day = 365, 
                                  within = TRUE,
                                  breeding_categories = c("C4", "C3", "C2", 
                                                          "C1", ""))

pres.blocks_sf <- right_join(blocks_sf, pres.blocks, by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = pres.blocks_sf, aes(fill = common_name)) +
  ggtitle("Blocks With Observations")
```

Map the blocks with breeding season observations
```{r}
# Pull out breeding season dates
breedates <- get_breeding_dates(species, day_year = TRUE)

summer.blocks <- blocks_observed_in(observations, start_day = breedates[1], 
                                    end_day = breedates[2], 
                                    within = TRUE,
                                    breeding_categories = c("C4", "C3", "C2", 
                                                            "C1", ""))

summer.blocks_sf <- right_join(blocks_sf, summer.blocks, 
                               by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = summer.blocks_sf, aes(fill = common_name)) +
  ggtitle("Blocks With Breeding Season Observations")
```

Map blocks with nonbreeding season observations
```{r}
winter.blocks <- blocks_observed_in(observations, start_day = breedates[1], 
                                    end_day = breedates[2], 
                                    within = FALSE,
                                    breeding_categories = c("C4", "C3", "C2", 
                                                            "C1", ""))

winter.blocks_sf <- right_join(blocks_sf, winter.blocks, 
                               by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = winter.blocks_sf, aes(fill = common_name)) +
  ggtitle("Blocks With Nonbreeding Season Observations")
```

Map blocks with confirmed or probable breeding observations from any day of the year.
```{r}
confirmed <- blocks_observed_in(observations, start_day = 1, 
                                end_day = 365, 
                                within = TRUE,
                                breeding_categories = c("C3", "C4"))

confirmed_sf <- right_join(blocks_sf, confirmed, 
                   by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = confirmed_sf, aes(fill = common_name)) +
  ggtitle("Blocks with Confirmed or Probable Breeding Records")
```

Map blocks with confirmed or probable breeding observations from within breeding safe dates.
```{r}
conf.prob <- blocks_observed_in(observations, start_day = breedates[1], 
                                end_day = breedates[2], 
                                within = TRUE,
                                breeding_categories = c("C3", "C4"))

conf.prob_sf <- right_join(blocks_sf, conf.prob, 
                           by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = confirmed_sf, aes(fill = common_name)) +
  ggtitle("Blocks with Confirmed or Probable Breeding Records Within Safe Dates")
```

Map blocks with confirmed or probable breeding observations from outside of breeding safe dates.
```{r}
conf.prob <- blocks_observed_in(observations, start_day = breedates[1], 
                                end_day = breedates[2], 
                                within = FALSE,
                                breeding_categories = c("C3", "C4"))

conf.prob_sf <- right_join(blocks_sf, conf.prob, 
                           by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = confirmed_sf, aes(fill = common_name)) +
  ggtitle("Blocks with Confirmed or Probable Breeding Records Outside Safe Dates")
```



```{r}
# Probable
probable <- blocks_observed_in(observations, start_day = breedates[1], 
                                    end_day = breedates[2],
                                    within = TRUE,
                                    breeding_categories = c("C3"))

probable_sf <- right_join(blocks_sf, probable, 
                   by = join_by("ID_BLOCK_CODE" == "atlas_block"))

# Plot the spatial data frame
ggplot() +
  geom_sf(data = blocks_sf) +
  geom_sf(data = probable_sf, aes(fill = common_name)) +
  ggtitle("Blocks with Probable Breeding")
```
# Tests

# Speed

# Bugs
present column isn't being renamed appropriately.