---
title: "dev_breeding_map"
author: "N.M. Tarr"
date: "2023-12-7"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Function
Uses the TMAP package to create an interactive map of the provided records.  Designed for use with output from get_breeding_records().  Hovering the mouse over the record symbol displays the sampling event identifier.  Clicking on symbols opens a small table showing records attributes.  The popup.vars argument controls what is included in those tables.  Including "URL" will provide a hyperlink to the checklist webpage on ebird.com.
```{r}
breeding_map
```

# Usage
When paired with get_breeding_records(), the function allows the user to map and explore the attributes of records.
```{r}
breeding_map("Orchard Oriole")
```

```{r}
breeding_map("Mississippi Kite")
```