---
title: "dev_breeding_map"
author: "N.M. Tarr"
date: "2023-12-7"
output:
  word_document: html
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Function
Uses the TMAP package to create an interactive map of the provided records.  Designed for use with output from get_breeding_records().  Hovering the mouse over the record symbol displays the sampling event identifier.  Clicking on symbols opens a small table showing records attributes.  The popup.vars argument controls what is included in those tables.  Including "URL" will provide a hyperlink to the checklist webpage on ebird.com.
```{r}
breeding_map
```

# Usage
When paired with get_breeding_records(), the function allows the user to map and explore the attributes of records.
```{r}
# Get breeding records
df <- get_breeding_records(behaviors = c("PE", "FS")) %>%
  to_EBD_format() %>%
  auk_unique()

# Make a map
breeding_map(records = df, popup.vars = c("URL", "common_name", "behavior_code"), 
             title = "Physiological Evidence or Carrying a Fecal Sac")
```

# Tests
The usage example shows a case where records are for two behavior codes, any species.  Try a case where records are from one species and one code: singing Turkey Vultures.
```{r}
species <- "Turkey Vulture"
records <- get_observations(species = species, EBD_fields_only = FALSE) %>%
  to_EBD_format() %>%
  auk_unique()

suspicious <- records %>% 
  filter(behavior_code == "S") %>%
  select(-c(state, state_code, country, bcr_code, usfws_code, iba_code, 
            locality_id, year, month, taxonomic_order, category))

breeding_map(suspicious, popup.vars = c("URL", "observation_date"), 
             title = "Singing Turkey Vultures")
```