---
title: "dev_records_as_sf"
author: "N.M. Tarr"
output:
  word_document: default
  html_document: default
---
```{r setup, include=TRUE, message=TRUE, warning=TRUE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```

# Purpose
This document details a function that creates a spatial representation of checklists or observations.  SHOULD A MAX EFFORT DISTANCE PARAMETER BE ADDED?
```{r}
records_as_sf_dev <- function(records_df, kind, method, fill_na_km = 0.1) {
  # Create new simple features (spatial data frame) of records (checklists or 
  #   observations)  Output can be plotted or used as input for other functions.
  # 
  #   Description: 
  #   Checklist records often need to be assigned geometries for visualization
  #   and spatial analyses, and different methods could be used.  Checklists
  #   can be represented as points or polygons, and polygons could be drawn as 
  #   buffers around the checklist coordinates (circles) or buffers drawn around
  #   checklist tracks.  Buffer length is meant to represent locational
  #   uncertainty (spatial precision) and can be approximated in different ways.
  #   Stationary or short lists should likely be buffered 100 m or more to 
  #   at least partially account for area surveyed.  Null effort_distance_km 
  #   values are filled with zero, which assumes those records are stationary 
  #   counts.
  #
  #   Parameters:
  #   records_df -- data frame of records with latitude, longitude, 
  #     checklists_id or sampling_event_identifier, atlas_block, protocol_type,
  #     and effort_distance_km columns.
  #   kind -- "checklists" or "observations" to identify what type of records are
  #     in the data frame.  Individual species data will be observations.
  #   method -- how to represent each record spatially.  Options are "points",
  #     "point-radius", and "buffered-tracks".
  #   fill_na_km -- NA values may exist in some records, which precludes creating
  #     a point-radius polygon for them. Enter a km distance to use as a 
  #     replacement.  This argument is inconsequential for the point method.
  #   
  #   Results:
  #   A spatial (simple features) data frame with columns for sampling_event_identifier or 
  #     sampling_event_identifier, atlas_block, protocol_type, 
  #     effort_distance_km, latitude, longitude and observation_count and
  #     breeding_code if the kind is observations.
  library(sf)
  
  if (kind == "checklists"){
    records_df <- records_df %>%
      select(sampling_event_identifier, atlas_block, protocol_type, effort_distance_km,
             latitude, longitude)
  } else {
    records_df <- records_df %>%
      select(sampling_event_identifier, atlas_block, protocol_type, 
             effort_distance_km, latitude, longitude, observation_count, 
             breeding_code)
  }
  
  # Make spatial frame
  checklists_sf <- records_df %>%
    st_as_sf(coords=c("longitude", "latitude"), crs=4326) %>%
    st_transform(6542)
  
  # Apply method
  if (method == "points") {
    checklists_sf <- checklists_sf
  }
    
  if (method == "point-radius") {
    checklists_sf <- checklists_sf %>%
      # Buffer coordinates
      replace_na(list(effort_distance_km=fill_na_km)) %>%
      mutate(buffer_length = (effort_distance_km)*1000) %>%
      mutate(footprint = st_buffer(geometry, buffer_length)) %>%
      select(-c(geometry)) %>%
      mutate(geometry = footprint) %>%
      st_set_geometry("geometry")
  }
    
  if (method == "buffer-tracks") {
    print("This method is currently unavailable until we get checklist tracks.")
  }
  
  return(checklists_sf)
}
```


# Usage
This demo requires the tidyverse packages.
```{r}
library(tidyverse)
```

Load the NCBA functions because this function relies upon the output from some of them.
```{r}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")
config <- "~/Documents/NCBA/Scripts/ncba_config.R"
```

Set the working directory to somewhere outside of the NCBA repository so that results are not saved in the repository.
```{r}
setwd("~/Documents/NCBA/species/")
```

Use the NCBA connection function to connect to the Atlas Cache (the mongodb).
```{r}
# connect to a specific collection (table)
connection <- connect_ncba_db(ncba_config = config, database = "ebd_mgmt", 
                              collection = "ebd")
```

ncba_functions has a function that returns a spatial data frame of NC counties in the 6542 CRS.  Get that for use later in maps.
```{r}
counties <- counties_NC()
```


## Observations (species-level records)
Identify a species to investigate.
```{r}
species <- "Virginia Rail"
```

Retrieve the records for the species from the Atlas Cache
```{r}
time1 <- proc.time()

nc_data <- get_observations(species, database = "AtlasCache", NCBA_only = TRUE,
                            EBD_fields_only = TRUE, ncba_config = config)

# format columns to the standard analysis format (ebd format)
records <- to_EBD_format(nc_data, drop=FALSE)

# Calculate processing time
mongotime <- proc.time() - time1

# Print number of records returned
print(paste("Records returned:", nrow(records)))
print(paste("Runtime: ", mongotime[["elapsed"]]))
```

Plot the coordinates
```{r}
coords <- records_as_sf_dev(records, kind="observations", method="points")

# Make a simple map
ggplot(data = counties) +
    geom_sf() +
    geom_sf(data = coords, size = 4, shape = 20, 
            crs = st_crs(6542), colour = "darkred") + 
    ggtitle(paste("All Record Coordinates for the", species), subtitle = "")
```

Plot the footprints.
```{r}
footprints <- records_as_sf_dev(records, kind="observations", 
                              method="point-radius")

# Make a simple map
ggplot(data = counties) +
    geom_sf() +
    geom_sf(data = footprints, size = 4, shape = 20, 
            crs = st_crs(6542), colour = "darkred", fill = NA) + 
    ggtitle(paste("All Record Footprints for the", species), subtitle = "")
```

Preview the result as a data frame
```{r}
print(head(footprints))
```

Plot just the records that confirm breeding
```{r}
# Get our list of confirmation codes from ncba_functions
conf_codes <- breeding_codes()["confirmed"][[1]]

# Select only those records that confirm
confirmed <- records %>% filter(breeding_code %in% conf_codes)

# Make it spatial
confirmed_sf <- records_as_sf_dev(confirmed, kind="observations", 
                                  method="points")

# Make a simple map
ggplot(data = counties) +
    geom_sf() +
    geom_sf(data = confirmed_sf, size = 4, shape = 20, 
            crs = st_crs(6542), colour = "darkred") + 
    ggtitle(paste("Coordinates for Confirmed Breeding by the", species), subtitle = "")
```

## Checklists
Retrieve a dataframe of checklists.  Note that this use depends on connection code from above.
```{r}
checklists <- get_checklists(ncba_config = config) %>% 
  to_EBD_format(drop=TRUE)
```

Plot the coordinates for all of the checklists from the Atlas Cache.
```{r}
# Make the data frame spatial
footprints <- records_as_sf_dev(checklists, kind="checklists", method="points")

# Make a simple map
ggplot(data = counties) +
       geom_sf() +
       geom_sf(data = footprints, size = 4, shape = 3, crs = st_crs(6542),
               colour = "darkred") + 
       ggtitle("All NC Atlas Cache Checklist Coordinates", subtitle = "")
```

