---
title: "dev_get_blocks"
author: "N.M. Tarr"
date: "2023-08-28"
output: word_document
---

```{r setup, include=TRUE, message=TRUE, warning=TRUE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
knitr::opts_knit$set(root.dir = '~/Documents/NCBA/species/')
```


# Purpose
This document details a function to retrieve information about NCBA blocks in the form of a data frame or a spatial data frame (simple feature).  The function also allows the user to pass a list of fields to return if they do not want all of the available fields.  Restricting the fields returned can drastically speed up the query because some fields are nested with a large amount of data.


# Function
```{r}
get_blocks
```


# Usage
This demo requires the tidyverse packages.
```{r}
library(tidyverse)
```

Load the NCBA functions because this function relies upon the output from some of them.
```{r}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")
ncba_config <- "~/Documents/NCBA/Scripts/ncba_config.R"
```

Retrieve a blocks data frame with all columns
```{r}
blocks <- get_blocks(ncba_config = ncba_config, spatial = FALSE)
print(head(blocks, 3))
```

Retrieve a blocks data frame with a subset of columns
```{r}
fields <- c("ID_BLOCK", "ECOREGION", "COUNTY")
blocks <- get_blocks(ncba_config = ncba_config, spatial = FALSE,
                         fields = fields)
print(head(blocks))
```

Retrieve a spatial data frame with a few columns
```{r}
fields <- c("ID_BLOCK", "ECOREGION")
blocks_sf <- get_blocks(ncba_config = ncba_config, spatial = TRUE,
                         fields = fields)
print(head(blocks_sf, 5))
```


# Tests
Plot a blocks spatial data frame.
```{r}
fields <- c("ID_BLOCK", "TYPE")
blocks_sf <- get_blocks(ncba_config = ncba_config, spatial = TRUE,
                         fields = fields)

# Plot the spatial data frame
ggplot(blocks_sf) +
  geom_sf()
```

Get a data frame with a subset of fields
```{r}
fields <- c("ID_BLOCK", "ECOREGION")
blocks <- get_blocks(ncba_config = ncba_config, spatial = FALSE,
                         fields = fields)
print(names(blocks))
```

Confirm that the default CRS of output is 4326
```{r}
fields <- c("ID_BLOCK")
blocks_sf <- get_blocks(ncba_config = ncba_config, spatial = TRUE,
                            fields = fields)
st_crs(blocks_sf)
```

Confirm that the CRS changes appropriately when requested.
```{r}
fields <- c("ID_BLOCK")
blocks_sf <- get_blocks(ncba_config = ncba_config, spatial = TRUE,
                            fields = fields, crs = 5070)
st_crs(blocks_sf)
```


# Speed
Describe runtimes for queries of a subset of fields
```{r}
fields <- c("ID_BLOCK", "ECOREGION", "COUNTY", "ID_WEB_BLOCKMAP")

#Run the function 30 times and record the runtime
time <- c()
for (i in 1:30) {
  time1 <- proc.time()
  get_blocks(ncba_config = ncba_config, fields = fields)
  time[i] <- proc.time() - time1
}

# Print the descriptive statistics
print(summary(time))
```

Describe runtimes for subset queries with geometries (spatial set to TRUE)
```{r}
fields <- c("ID_BLOCK", "ECOREGION", "COUNTY", "ID_WEB_BLOCKMAP")

# Run the function 30 times and record the runtime
time <- c()
for (i in 1:30) {
  time1 <- proc.time()
  get_blocks(ncba_config = ncba_config, fields = fields, spatial = TRUE)
  time[i] <- proc.time() - time1
}

# Print the descriptive statistics
print(summary(time))
```

Describe runtimes for query of all fields
```{r}
# Run the function 5 times and record the runtime
time <- c()
for (i in 1:5) {
  time1 <- proc.time()
  get_blocks(ncba_config = ncba_config, fields = NULL)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```