---
title: "dev_block_predicted_spp_all_blocks"
author: "N.M. Tarr, S.K. Anderson"
date: "2023-10-11"
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# setwd("~/Code/NCBA/resources")
if(!require(here)) install.packages(
  "here", repos = "http://cran.us.r-project.org")
source(here("resources", "ncba_functions.R"))

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


# Purpose
This document details a function to retrieve a list of species that were predicted to occur in a block by either the USGS National Gap Analysis Project (2001) or from analyzing eBird data.   

# Function
```{r}
block_predicted_spp
```

# Usage
First retrieve list of priority blocks.
```{r}
# block <- "35079H1SE"
blocks <- get_blocks(
  spatial = FALSE,
  fields = c(
    "ID_NCBA_BLOCK",
    "ID_BLOCK_CODE",
    "ECOREGION",
    "COUNTY"
  ),
  priority_only = TRUE
  )

```

Loop through blocks.
```{r}


for( i in rownames(blocks) ){
  print( blocks[i,"ID_NCBA_BLOCK"])

  GAP_predictions <- block_predicted_spp(
    block = blocks[ i, "ID_BLOCK_CODE"],
    source = "GAP"
    )

  print(GAP_predictions$summer)
  
  # for testing
  if ( i > 5 ) {
    break
    }
}
```

Retrieve the predicted species lists from GAP.
```{r}

for( i in rownames(blocks) )
  print( blocks[i,"ID_NCBA_BLOCK"])


GAP_predictions <- block_predicted_spp(block = block, source = "GAP")
```

Print the GAP prediction for summer
```{r}
print(GAP_predictions$summer)
```

Print the GAP prediction for winter
```{r}
print(GAP_predictions$winter)
```

Retrieve the predicted species lists from eBird.
```{r}
eBird_predictions <- block_predicted_spp(block = block, source = "eBird")
```

Print the eBird prediction for summer.
```{r}
print(eBird_predictions$summer)
```

Print the eBird prediction for winter
```{r}
print(eBird_predictions$winter)
```


# Speed
Run the function several times to get descriptive statistics.
```{r}
# Run the function 10 times and record the runtime
time <- c()
for (i in 1:10) {
  time1 <- proc.time()
  block_predicted_spp(block = block, source = "eBird")
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```