---
title: "dev_highest_category"
author: "N.M. Tarr"
date: "2023-09-20"
output:
  word_document: default
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Function
This function creates and returns a data frame of the highest breeding code reported for in each atlas block for a species.
```{r}
highest_category
```

# Usage
Get a data frame of blocks with the highest reported code for the species.
```{r}
species <- "Yellow-breasted Chat"

highest.sf.1 <- highest_category(species, dataframe = NULL)
```

Plot the spatial data frame.
```{r}
ggplot() +
  geom_sf(data = highest.sf.1, aes(fill = highest_category)) + 
  ggtitle("Highest Reported Breeding Category")
```

Preview the output data frame.
```{r}
print(head(highest.sf.1))
```


Run the function utilizing the data frame parameter.
```{r}
# Get an observations data frame
observations <- get_observations(species = "Magnolia Warbler", 
                        database = "AtlasCache",
                        NCBA_only = TRUE,
                        EBD_fields_only = TRUE,
                        fields = NULL) %>%
  to_EBD_format(drop = TRUE)

# Get the highest codes
highest.sf.2 <- highest_category(species, dataframe = observations)
```

Plot the spatial data frame
```{r}
ggplot() +
  geom_sf(data = highest.sf.2, aes(fill = highest_category)) + 
  ggtitle("Highest Reported Breeding Code")
```

Preview the output data frame.
```{r}
print(head(highest.sf.2))
```

# Tests
How should this function be tested?


# Speed (in seconds)
With the acquisition of a data frame as part of the process, for a data-poor species.
```{r}
species <- "Willow Flycatcher"

# Run the function 3 times and record the runtime
time <- c()
for (i in 1:3) {
  time1 <- proc.time()
  highest_category(species, dataframe = NULL)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```

With the acquisition of a data frame as part of the process, for a data-rich species.
```{r}
species <- "Red-eyed Vireo"

# Run the function 2 times and record the runtime
time <- c()
for (i in 1:2) {
  time1 <- proc.time()
  highest_category(species, dataframe = NULL)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```

Without the acquisition of a data frame as part of the process.
```{r}
# Run the function 3 times and record the runtime
time <- c()
for (i in 1:3) {
  time1 <- proc.time()
  highest_category(species, dataframe = observations)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```