---
title: "dev_breeding_map"
author: "N.M. Tarr"
date: "2023-12-7"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Function
Uses the TMAP package to create an interactive map of the provided records.  Designed for use with output from get_breeding_records().  Hovering the mouse over the record symbol displays the sampling event identifier.  Clicking on symbols opens a small table showing records attributes.  The popup.vars argument controls what is included in those tables.  Including "URL" will provide a hyperlink to the checklist webpage on eBird.com. 
```{r}
breeding_map <- function(species) {
  # Makes an interactive map of records color coded by breeding category
  
  #
  # Description:
  # Gets observations for a species and maps them atop a counties layer in an
  #   interactive map.  Clicking on record symbols reveals a hyperlink to the 
  #   checklist webpage on ebird.com.
  #
  # Arguments:
  # species -- common name
  
  # Get the records
  records <- get_observations(species = species, EBD_fields_only = FALSE) %>%
    to_EBD_format() %>%
    auk_unique()
  
  # Make spatial data frames of all records
  records_sf <- records_as_sf(records, kind = "observations", method = "points") %>%
    right_join(records, by = "sampling_event_identifier") %>%
    filter(breeding_category != "")
  
  # Add a column with code to open the webpage for each checklist
  records_sf$front <- '<a href = https://ebird.org/checklist/'
  
  # Strip off any excess identifiers (group checklists produce "S104604778,S104604779")
  records_sf$sampling_event_identifier <- lapply(strsplit(records_sf$sampling_event_identifier, split = ","), function(l) l[[1]])
  records_sf$URL <- with(records_sf, paste0(front, sampling_event_identifier,
                                            ">visit</a>"))
  
  # Draw the map
  tmap_mode("view") 
  tm_shape(shp = counties_NC(), name = "counties") + tm_borders() +
    tm_shape(shp = records_sf, name = "observations") + 
    tm_dots(interactive = TRUE, popup.vars = c("URL", "observation_date", "behavior_code"), col = "breeding_category",
            popup.format = list(html.escape = F), border.alpha = 0#,
            #palette = c("yellow", "lightgreen", "darkgreen", "purple"),
            #labels = c("observed", "possible", "probable", "confirmed")
            ) +
    tm_layout(title = species)
}

```

# Usage
When paired with get_breeding_records(), the function allows the user to map and explore the attributes of records.
```{r}
breeding_map("Orchard Oriole")
```

```{r}
breeding_map("Mississippi Kite")
```

# Tests
View a case where not all breeding categories are represented.
```{r}
breeding_map("Worm-eating Warbler")
```