---
title: "dev_get_checklists"
author: "N.M. Tarr"
date: "2023-08-28"
output: html_document
---

```{r setup, include=TRUE, message=TRUE, warning=TRUE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
knitr::opts_knit$set(root.dir = '~/Documents/NCBA/species/')
```

# Function
```{r}
get_checklists_dev <- function(dataset = "AtlasCache", drop_columns = TRUE, 
                               ncba_config){
  # Get a data frame of checklists from the AtlasCache
  # 
  # Parameters:
  # dataset -- either "EBD" for a downloaded eBird dataset or "AtlasCache" for
  #   the NCBA mongodb.
  # drop_columns -- whether to drop non-eBird columns from the output data frame.
  #   TRUE or FALSE and defaults to TRUE.
  # ncba_config -- config file with NCBA MongoDB username and password

  library(tidyverse)
  
  # Connect to the NCBA database
  connection <- connect_ncba_db(ncba_config, "ebd_mgmt", "ebd")
  
  # Define a query
  query <- '{}'
  
  # Define a filter that excludes the observation column
  filter <- '{"OBSERVATIONS":0}'
  
  # Retrieve the checklists
  checklists <- connection$find(query = query, fields = filter)
  
  # Make column names lower case to match eBird data download format
  colnames(checklists) <- tolower(colnames(checklists))
  
  # Drop columns not in eBird sampling database.
  colnames(checklists)[1] <- c("checklist_id")
  
  if (drop_ncba_col == TRUE) {checklists %>% select(
    c(checklist_id, last_edited_date, county, county_code,
      iba_code, bcr_code, usfws_code, atlas_block,
      locality, locality_id, locality_type, latitude,
      longitude, observation_date, time_observations_started,
      observer_id, sampling_event_identifier, protocol_type,
      protocol_code, project_code, duration_minutes, 
      effort_distance_km, effort_area_ha, number_observers,
      all_species_reported, group_identifier, trip_comments))}
  else {checklists <- checklists}
  
  # Coerce data types of some columns to match eBird database format
  checklists <- transform(checklists, bcr_code = as.integer(bcr_code),
                          duration_minutes = as.integer(duration_minutes),
                          all_species_reported = as.logical(all_species_reported),
                          observation_date = as.Date(observation_date))
}
```

# Usage
This demo requires the tidyverse packages.
```{r}
library(tidyverse)
```

Load the NCBA functions because this function relies upon the output from some of them.
```{r}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")
config <- "~/Documents/NCBA/Scripts/ncba_config.R"
```

Retrieve observation functions with the function
```{r}
get_checklists_dev(dataset = "AtlasCache", 
                   drop_columns = FALSE, ncba_config = config)
```

# Tests

# Speed
```{r}
# # Run the function 30 times and record the runtime
# time <- c()
# for (i in 1:30) {
#   time1 <- proc.time()
#   get_observations_dev(species = "Willow Flycatcher", dataset = "AtlasCache", 
#                        drop_columns = FALSE)
#   time[i] <- proc.time() - time1
# }
# 
# # Print the descriptive statistics
# print(summary(time))