---
title: "dev_get_checklists"
author: "N.M. Tarr"
date: "2023-10-05
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Purpose
This document details the primary function for retrieving data at the checklist level, whether from AtlasCache or EBD.  The function "get_observations" provides for species observation records.It offers a few parameters that enable some filtering of records, which is provided in hopes of increasing query speed.  However, the function can still be somewhat slow.  Note that parameter defaults exist that affect the results even if the parameter is not explicitly called.  The project, block, and observer parameters enable the exclusion of all records that are not from a specified block, observer, and project.  EBD_fields_only and fields parameters allow the exclusion of all but specified fields (fields) and exclusion of all fields that are not present in the EBD (EBD_fields_only).

# Function
```{r}
get_checklists
```


# Usage
Retrieve all checklists from the Atlas Cache, but exclude non-EBD fields and print the number of rows in the data frame to see how many checklist records are available.
```{r}
checklists <- get_checklists(database = "AtlasCache", project = NULL, 
                             EBD_fields_only = TRUE)
print(nrow(checklists))
```

Retrieve all checklists from an EBD download for a single block.
```{r}
checklists.EBD <- get_checklists(database = "EBD", block = "35079E2SW")
print(head(checklists.EBD))
```

Get checklists from specific observer, block, and project.
```{r}
specific <- get_checklists(database = "AtlasCache", project = "EBIRD_ATL_NC", 
                           observer = "obsr2744483", block = "35079E2SW")
print(specific)
```

Retrieve only a few columns
```{r}
few_cols <- get_checklists(database = "AtlasCache", project = "EBIRD_ATL_NC", 
                           observer = "obsr2744483", block = "35079E2SW",
                           fields = c("LOCALITY_ID", "BCR_CODE", 
                                      "PROTOCOL_TYPE"))
print(few_cols)
```

Get all the checklists for an observer.
```{r}
ob <- get_checklists(observer = "obsr447159", project = NULL)
print(head(ob))
```


# Tests

The default data frame returned from the AtlasCache should not include the OBSERVATIONS field.  This chunk should return FALSE to pass the test.
```{r}
checklists <- get_checklists(database = "AtlasCache")

print("OBSERVATIONS" %in% names(checklists))
```

Records from the Atlas Cache should also not include the nonEBD fields when EBD_fields_only is set to TRUE, which is the default.  Thus to pass this test, the only value returned by this chunk should be FALSE.
```{r}
print(unique(nonEBD_fields() %in% names(checklists)))
```

On the other hand, if EBD_fields_only is set to FALSE, then non-EBD fields should be present in a returned data frame and the test from above should return TRUE (and possibly FALSE as well).
```{r}
checklists <- get_checklists(database = "AtlasCache",
                                 EBD_fields_only = FALSE)

print(unique(nonEBD_fields() %in% names(checklists)))
```

There also shouldn't be an OBSERVATIONS column in the data frame.  This chunk should return FALSE to pass the test.
```{r}
print("OBSERVATIONS" %in% names(checklists))
```

Utilizing the fields parameters should limit the fields returned.  Thus, the column names from the output data frame should be the same as the fields argument and this chunk should return TRUE to pass the test.
```{r}
# Make a list of fields for a test.  Choose some from EBD and others from NCBA-specific fields.
testfields <- c("_id", "LOCALITY_ID", "SAMPLING_EVENT_IDENTIFIER", 
                "NCBA_SEASON", "GEOM") 

checklists2 <- get_checklists(database = "AtlasCache", fields = testfields)

unique(sort(names(checklists2)) == sort(testfields))
```

EBird checklists that were submitted with the NCBA portal are assigned the value "EBIRD_ATL_NC" in the PROJECT_CODE field.  Therefore, using the NCBA_only parameter should return a data frame with no values other than EBIRD_ATL_NC in the PROJECT_CODE field.  This test is passed if the chunk returns "EBIRD_ATL_NC".
```{r}
checklists3 <- get_checklists(database = "AtlasCache", project = "EBIRD_ATL_NC")
print(unique(checklists3$PROJECT_CODE))
```

Perform the same test, but with the EBD database.
```{r}
checklists4 <- get_checklists(database = "EBD", project = "EBIRD_ATL_NC")
print(unique(checklists4$project_code))
```

Limiting on observer and project should only return records with the requested values.  This chunk should return TRUE twice and never FALSE to pass the test.
```{r}
observer <- "obsr547676"
project <- "EBIRD"

multi <- get_checklists(database = "AtlasCache", project = project, 
                        observer = observer, block = NULL)

print(unique(multi$PROJECT_CODE) == project)
print(unique(multi$OBSERVER_ID) == observer)
```

If a block is specified, then only records from that block should be returned.  This chunk should return TRUE to pass.
```{r}
block <- "35079E2SW"
checklists5 <- get_checklists(block = block)
print(unique(checklists5$ATLAS_BLOCK) == block)
```


# Speed (in seconds)
Describe runtimes for queries of the Atlas Cache for NCBA records
```{r}
# Run the function 3 times and record the runtime
time <- c()
for (i in 1:3) {
  time1 <- proc.time()
  get_checklists()
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```


Measure the runtime of an EBD query, which tend to be very slow, so only do that one
once.
```{r}
# time1 <- proc.time()
# get_checklists(database = "EBD")
# print(proc.time() - time1)
```




