---
title: "dev_summarize_duration"
author: "N.M. Tarr"
date: "2023-10-10"
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")
library(auk)

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Purpose
This document details a function for summarizing hours of effort within a block.  Options are provided to choose whether to summarize by month or year (or both) and to customize a period of the year for the summary.      

# Function
```{r}
summarize_duration
```

# Usage
First, identify a block of interest.
```{r}
block <- "35079H1SE"
```

To summarize within a time period specify the start and end days as numbered days of the year and use the "within" option.  Records from between the start and end date will be used.
```{r}
start_day <- 70
end_day <- 230

summer.summary <- summarize_duration(block = block, summarize_by = "year", 
                                     start_day = start_day,
                                     end_day = end_day, 
                                     within = TRUE)

print(summer.summary)
```

To summarize dates outside of a time period (e.g., nonbreeding season), use the within option.
```{r}
winter.summary <- summarize_duration(block = block, summarize_by = "year", 
                                     start_day = start_day,
                                     end_day = end_day, 
                                     within = FALSE)

print(winter.summary)
```

Summaries can be based on month...
```{r}
winter.summary2 <- summarize_duration(block = block, summarize_by = "month", 
                                     start_day = start_day,
                                     end_day = end_day, 
                                     within = FALSE)

print(winter.summary2)
```

Or year and month...
```{r}
winter.summary3 <- summarize_duration(block = block, summarize_by = "year-month", 
                                      start_day = start_day,
                                      end_day = end_day, 
                                      within = FALSE)

print(winter.summary3)
```

If a table of winter and summer together were desired, merely add a column to each and then use "rbind" to combine them.
```{r}
summer.summary$season <- "summer"
winter.summary$season <- "winter"
all.summary <- rbind(summer.summary, winter.summary)
print(arrange(all.summary, year))
```


# Speed
Run the function several times to get descriptive statistics.
```{r}
# Run the function 3 times and record the runtime
time <- c()
for (i in 1:3) {
  time1 <- proc.time()
  summarize_duration(block = block, summarize_by = "year", 
                     start_day = start_day, end_day = end_day,
                     within = FALSE)  
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```