---
title: "dev_get_checklists"
author: "N.M. Tarr"
date: "2023-08-28"
output: html_document
---

```{r setup, include=TRUE, message=TRUE, warning=TRUE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Function
```{r}
get_checklists
```

# Usage

Retrieve all checklists from the Atlas Cache.
```{r}
checklists <- get_checklists(database = "AtlasCache")
print(head(checklists))
```

# Tests

The default data frame returned from the AtlasCache should not include the OBSERVATIONS field.  This chunk should return FALSE to pass the test.
```{r}
checklists <- get_checklists(database = "AtlasCache")

print("OBSERVATIONS" %in% names(checklists))
```

Records from the Atlas Cache should also not include the nonEBD fields when EBD_fields_only is set to TRUE, which is the default.  Thus, the only value returned by this chunk should be FALSE to pass this test.
```{r}
print(unique(nonEBD_fields() %in% names(checklists)))
```

On the other hand, if EBD_fields_only is set to FALSE, then non-EBD fields should be present in a returned data frame and the test from above should return TRUE in this chunk.
```{r}
checklists <- get_checklists(database = "AtlasCache",
                                 EBD_fields_only = FALSE)

print(unique(nonEBD_fields() %in% names(checklists)))
```

There also shouldn't be an OBSERVATIONS column in the data frame.  This chunk should return FALSE to pass the test
```{r}
print("OBSERVATIONS" %in% names(checklists))
```


Utilizing the fields parameters should limit the fields returned.  Thus, the column names from the output data frame should be the same as the fields argument and this chunk should return TRUE to pass the test.
```{r}
# Make a list of fields for a test.  Choose some from EBD and others from NCBA-specific fields.
testfields <- c("_id", "LOCALITY_ID", "SAMPLING_EVENT_IDENTIFIER", 
                "NCBA_SEASON", "GEOM") 

checklists2 <- get_checklists(database = "AtlasCache", fields = testfields)

unique(sort(names(checklists2)) == sort(testfields))
```

EBird checklists that were submitted with the NCBA portal are assigned the value "EBIRD_ATL_NC" in the PROJECT_CODE field.  Therefore, using the NCBA_only parameter should return a data frame with no values other than EBIRD_ATL_NC in the PROJECT_CODE field.  This test is passed if the chunk returns "EBIRD_ATL_NC".
```{r}
checklists3 <- get_checklists(database = "AtlasCache", NCBA_only = TRUE)
print(unique(checklists3$PROJECT_CODE))
```

Perform the same test, but with the EBD database.  WARNING: this is very slow.
```{r}
# checklists4 <- get_checklists(database = "EBD", NCBA_only = TRUE)
# print(unique(checklists4$project_code))
```


# Speed (in seconds)
Describe runtimes for queries of the Atlas Cache with ncba_only = TRUE
```{r}
# Run the function 5 times and record the runtime
time <- c()
for (i in 1:5) {
  time1 <- proc.time()
  get_checklists(database = "AtlasCache", NCBA_only = TRUE)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```


Measure the runtime of an EBD query, which tend to be very slow, so only do that one
once.
```{r}
# time1 <- proc.time()
# get_checklists(database = "EBD", NCBA_only = TRUE)
# print(proc.time() - time1)
```




