---
title: "dev_calculate_breeding_dates"
author: "N.M. Tarr"
date: "2023-09-13"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)

print(paste("The document working directory is: ", getwd()))
```

# Purpose
This document defines and demonstrates a function that computes breeding season start and end dates based upon NCBA data.  Results are computed statewide, as well as by ecoregion.


# Function
Note: I explored using the quantiles function to calculate directly from date type values, but the results did not seem correct and were possible biased by over a month.
```{r}
calculate_breeding_dates
```


# Usage
Running the function returns a list of dates.
```{r}
species <- "Yellow-throated Vireo"
quantiles <- c(.05, .95)

dates <- calculate_breeding_dates(species = species, basis = ("confirmed"), 
                                  quantiles = quantiles, year = 2023,
                                  year_day = FALSE)
print(dates)
```

Running the function with year_day set to TRUE returns a list of days of the year.
```{r}
dates <- calculate_breeding_dates(species = species, basis = ("confirmed"), 
                                  quantiles = quantiles, year = 2023,
                                  year_day = TRUE)
print(dates)
```


# Tests
Visually inspect whether the provided dates generally align with the breeding boxplot.
```{r}
species <- "Least Tern"

print(calculate_breeding_dates(species = species, basis = ("confirmed"), 
                               quantiles = quantiles, year = 2023,
                               year_day = FALSE))
```


```{r}
quantiles <- c(0.25, 0.75)
observations <- get_observations(species, database = "AtlasCache", 
                                 NCBA_only = TRUE, 
                                 EBD_fields_only = FALSE,
                                 fields = NULL) %>%
  to_EBD_format(drop=FALSE)

breeding_boxplot(species = species, data = observations, type = "ecoregional",
                 lump = breeding_codes(lumped = TRUE))
```

```{r}
breeding_boxplot(species = species, data = observations, type = "interactive",
                 lump = breeding_codes(lumped = TRUE))
```


# Speed
The calculations performed by the function require the retrieval of all records for a species, thus the runtime depends upon the rarity of the species.  Test a data-poor species.
```{r}
basis = c("confirmed", "probable")

# Run the function 5 times and record the runtime
time <- c()
for (i in 1:5) {
  time1 <- proc.time()
  calculate_breeding_dates(species = "Willow Flycatcher", basis = basis, 
                 quantiles = quantiles, year = 2023)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```


Test a data-rich species.
```{r}
# Run the function 3 times and record the runtime
time <- c()
for (i in 1:3) {
  time1 <- proc.time()
  calculate_breeding_dates(species = "Tufted Titmouse", basis = basis, 
                 quantiles = quantiles, year = 2023)
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```
