---
title: "dev_get_breeding_dates"
author: "N.M. Tarr"
date: "2023-09-19"
output:
  html_document: default
  word_document: default
---

```{r setup, include=TRUE, message=TRUE, warning=TRUE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)

print(paste("The document working directory is: ", getwd()))
```

# Purpose
This document defines and demonstrates a function that gets breeding season start and end dates from the AtlasCache.


# Function
Note: I explored using the quantiles function to calculate directly from date type values, but the results did not seem correct and were possible biased by over a month.
```{r}
get_breeding_dates
```


# Usage
Pick a species
```{r}
species <- "Orchard Oriole"
```

# Get the dates in month-day format
```{r}
dates <- get_breeding_dates(species = species)
print(dates)
```

# Get the dates in day of year format
```{r}
dates <- get_breeding_dates(species = species, day_year = TRUE)
print(dates)
```

# Tests
If this chunk returns TRUE, then it passes the test.
```{r}
# Enter Black Vulture dates after viewing them in MondoDB Compass
B_SAFE_START_DATE_BLVU <- "05-01"
B_SAFE_END_DATE_BLVU <- "07-31"

# Retrieve the dates with the function
dates <- get_breeding_dates(species = "Black Vulture")
print(dates[1] == B_SAFE_START_DATE_BLVU)
print(dates[2] == B_SAFE_END_DATE_BLVU)
```


# Speed
Run the function 30 times and record the runtime
```{r}
time <- c()
for (i in 1:30) {
  time1 <- proc.time()
  get_breeding_dates(species = "Willow Flycatcher")
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```