---
title: "dev_observer_complete_by_priority"
author: "N.M. Tarr"
date: "2023-10-06"
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# setwd("~/Code/NCBA/resources")
if(!require(here)) install.packages(
  "here", repos = "http://cran.us.r-project.org")
source(here("resources", "ncba_functions.R"))

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(auk)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Purpose
This document contains descriptions and tests of a function that provides tables summarizing how many of their checklists were complete by block type.  The summary can be presented in terms of the number of checklists, or the number of blocks with checklists by the observer.


# Function
```{r}
observer_complete_by_priority
```


# Usage
Get a table that summarizes the number of species that the observer reported, broken out by breeding category and atlas block type (priority or non-priority).
```{r}
table1 <- observer_complete_by_priority(observer = "obsr1000095",
                                        data = "checklists")

print(table1)
```

Get a table that summarizes the number of blocks from which the observer reported each breeding category, broken out by block type (priority or non-priority).
```{r}
table2 <- observer_complete_by_priority(observer = "obsr1000095",
                                        data = "blocks")

print(table2)
```


# Tests
The first sections of code in this function rely upon functions that were tested elsewhere, so only the latter portions of code need to be tested.  Test the function by verifying individual cell values from the output table with different methods than those used within the function.

Get a data frame with the same code as used within the function to test against. 
```{r}
library(auk)

observer = "obsr1000095"

# Get a data frame of atlas blocks with the priority column
blocks <- get_blocks(spatial = FALSE,
                     fields = c("ID_BLOCK_CODE", "PRIORITY"))

# Get a data frame of the observer's checklists, join get PRIORITY
checklists <- get_checklists(observer = observer) %>%
    to_EBD_format() %>%
    auk_unique(checklists_only = TRUE)

# Drop excess columns in checklists and join to get PRIORITY
checks0 <- checklists %>%
    select(c("atlas_block", "all_species_reported", "sampling_event_identifier")) %>%
    left_join(blocks, by = join_by("atlas_block" == "ID_BLOCK_CODE")) %>%
    data.frame()

# Replace the values in PRIORITY.
checks0$PRIORITY <- ifelse(checks0$PRIORITY == 1, "priority", "non-priority")

# Replace the values in all_species_reported.
checks0$all_species_reported <- ifelse(checks0$all_species_reported == TRUE, 
                                       "complete", "incomplete")
```

Get a table with checklist count as the data and extract the number of complete checklists in priority blocks.
```{r}
check.cnt <- observer_complete_by_priority(observer = observer, data = "checklists")

complete.priority <- check.cnt[1,2]

print(check.cnt)
```

Now find the number of complete checklists in priority blocks with a different method.  The test is passed if the chunk returns TRUE.
```{r}
tv0 <- checks0 %>%
  filter(PRIORITY == "priority", all_species_reported == "complete") %>%
  select(c("sampling_event_identifier"))
tv1 <- nrow(unique(tv0))

print(tv1 == complete.priority)
```

Do a similar test but with the number of blocks as the data type.
```{r}
block.cnt <- observer_complete_by_priority(observer = observer, data = "blocks")

complete.priority <- block.cnt[1,2]

print(block.cnt)
```

The test is passed if the chunk returns TRUE.
```{r}
tv2 <- checks0 %>%
  filter(PRIORITY == "priority", all_species_reported == "complete") %>%
  select(c("atlas_block"))
tv3 <- nrow(unique(tv2))

print(tv3 == complete.priority)
```


# Speed
Run the function several times to get descriptive statistics.
```{r}
# Run the function 3 times and record the runtime
time <- c()
for (i in 1:3) {
  time1 <- proc.time()
  observer_complete_by_priority(observer = observer, data = "checklists")
  t <- proc.time() - time1
  time[i] <- t["elapsed"]
}

# Print the descriptive statistics
print(summary(time))
```