---
title: "eBird Effort in North Carolina"
author: "N.M. Tarr"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  rmdformats::downcute:
  html_document:
    df_print: paged
    code_folding: hide
---
# Purpose
This notebook summarizes eBird sampling (effort) within North Carolina for the past 5 years.  Sampling data can be downloaded from the eBird website and queried with eBird's R package named "auk".  The script named "filter_eBird_sampling.R", which is saved in the NCBA github repo, performs the filtering that supports this notebook.  The notebook can be updated and revised periodically.

The records available from the sampling data set are for individual, unique checklists and several attributes are included for each.  This notebook summarizes the available attribute values to learn as much as possible about the recent eBird effort to date within North Carolina.  The available checklist attributes are:

checklist_id, last_edited_date, county, county_code, iba_code, bcr_code, 
usfws_code, atlas_block, locality, locality_id, locality_type, latitude,
longitude, observation_date, time_observations_started, observer_id, 
sampling_event_identifier, protocol_type, protocol_code, project_code, 
duration_minutes, effort_distance_km, effort_area_ha, number_observers,
all_species_reported, group_identifier, trip_comments 

```{r set_paths}
blocks_path <- "~/ENTER/YOUR/PATH/ncba_blocks.shp"
# Reference the filtered checklist output file
filtered_checklists <- "~/ENTER/YOUR/PATH/filtered_checklists.txt"
```

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE)

library(auk)
library(tidyverse)
library(ggplot2)
theme_set(theme_bw())
library(sf)
library(RColorBrewer)
library(maps)
library(hms)
library(lubridate)
library(tmap)
starttime <- Sys.time()

comma <- function(x) format(x, digits=2, big.mark=",")
```

```{r block_data}
# Read in NCBA block spatial data frame
blocks_sf <- st_read(blocks_path) %>% st_transform(6542)
st_crs(blocks_sf)
```

```{r checklists}
# Read in the effort data (filtered checklist table)
checklists <- read_sampling(filtered_checklists)

# Further filtering of the sampling data can occur here.  For example, we may
# want to remove data from certain projects.
```

```{r counties_data}
# Read in county spatial data frame
counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE)) %>%
  subset(grepl("north carolina", ID)) %>%
  mutate(county = str_to_title(str_replace(ID, "north carolina,", ""))) %>%
  select(-c(ID))
# Reproject?
```

# General

* There are currently `r comma(length(unique(checklists$checklist_id)))` checklists from North Carolina in eBird.

# Week
```{r}
by_week <- checklists %>%
  mutate(week=week(date(observation_date))) %>%
  group_by(week) %>%
  summarize(count=n())

ggplot(data=by_week) +
  geom_line(mapping=aes(y=count, x=week), show.legend=TRUE, color="orange") + 
  labs(title="EBirder activity varies greatly throughout the year",
       caption="Checklists from before 2021 are not included") +
  ylab("total number of checklists") +
  scale_x_continuous(limits=c(0,52), breaks=seq(0,52,by=4)) +
  scale_y_continuous(breaks=seq(0,30000,5000))
rm(by_week)
```


# Locality
```{r group_by_locality}
by_locality <- checklists %>%
  group_by(locality, locality_id) %>%
  summarize(count = n())
```

* Many checklists are revisits to previously surveyed locations.  The ratio of locations that were visited once to the total number of checklists is `r round(nrow(filter(by_locality, count==1))  / length(unique(checklists$checklist_id)), digits=2)`.

* Most locations have only been sampled once.  However, some locations have been visited a large number of times.

```{r locality_boxplot}
ggplot(data=by_locality) +
  geom_boxplot(mapping=aes(y=count, x=""), color="darkblue", 
               outlier.colour="orange", show.legend=TRUE) + 
  coord_flip() + 
  labs(title="Some locations have been very heavily sampled",
       caption="Checklists from before 2021 are not included") +
  ylab("Checklists (n)") + 
  scale_y_continuous(n.breaks=12)
```

```{r locality_pie}
labels <- c("1", "2:10", "11:20", "21:5,000")
locality_bins <- table(cut(by_locality$count, breaks=c(0,1,10,20,5000),
                           labels=labels)) %>%
                  data.frame()
colnames(locality_bins) = c("times_surveyed", "value")

ggplot(data=locality_bins, aes(x="", y=value, fill=times_surveyed)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_viridis_d(alpha = 1, option="D") +
  theme_void() + 
  labs(title="Most locations have only been sampled once",
       caption="Checklists from before 2021 are not represented")
```

```{r locality_type_pie}
# Summarize whether checklists are for hotspots, personal locations, etc.
by_locality_type <- checklists %>%
  group_by(locality_type) %>%
  summarize(
    count = n())

# Print pie chart
ggplot(data=by_locality_type, aes(x="", y=count, fill=locality_type)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_viridis_d(alpha = 1, option="D") +
  theme_void() + 
  labs(title="Many checklists are assigned to hotspots, but the majority are not",
       caption="Checklists from before 2021 are not represented")

# Print table
knitr::kable(by_locality_type, caption="Number of checklists per locality type")
```

## Most Visited Locations Table
```{r most_visited_locations}
most_visited <- by_locality %>%
  arrange(by=desc(count)) %>%
  filter(count > 99)

knitr::kable(most_visited, caption="Locations visited 100 or more times.")
rm(by_locality, by_locality_type)
```

# Spatial Distribution
```{r xy_locations}
ggplot(data=checklists) +
  geom_point(mapping=aes(y=latitude, x=longitude), color="darkgreen",
             shape=3) + 
  labs(title="Checklist locations are distributed throughout NC, but coverage is incomplete",
       caption="Checklists from before 2021 were not included in this summary") +
  ylab("latitude") + 
  xlab("longitude")
```

## Checklists per County
```{r per_county_bwplot}
by_county <- checklists %>%
  group_by(county) %>%
  summarize(
    count = n(),
  ) %>%
  arrange(county) 

# We will need median # checklists for below
median.checklists <- toString(comma(summary(by_county$count)[[3]]))

# Make a boxplot of checklists per county
ggplot(data=by_county) +
  geom_boxplot(mapping=aes(y=count, x=""), color="darkred", 
               outlier.colour="orange", show.legend=TRUE) + 
  coord_flip() + 
  labs(title="A handful of counties have many more checklists than others",
       subtitle=str_c("The median number of checklists in a county was ",
                      median.checklists),
       caption="Checklists from before 2021 are not included") +
  ylab("Checklists (n)") + 
  scale_y_continuous(n.breaks=12)
```

Between `r comma(summary(by_county$count)[[1]])` and `r comma(summary(by_county$count)[[6]])` checklists are available for counties in North Carolina, and the median number of checklists is `r comma(summary(by_county$count)[[3]])`.

```{r per_county_map}
# Join to get the checklist count
counties.checklists <- counties %>%
  left_join(by_county, by = "county")

# Pull out just the column of interest here ("count") and plot it.
county.count <- select(counties.checklists, count, geom)
ggplot(data=county.count) +
  geom_sf(data=counties.checklists, aes(fill=count)) +
  scale_fill_viridis_c(trans = "sqrt", alpha = 1, option="E") +
  labs(title="Large variation exists in the number of checklists among counties",
       caption="Checklists from before 2021 are not included")
```

### Top 10 Counties
```{r top_county_table}
knitr::kable(head(arrange(by_county, desc(count)), n=10), caption="Top 10 Counties")
```

```{r top_county_map}
# Add a column to record whether it's a "top" or "bottom" county as far as count goes.
tops <- counties.checklists %>% arrange(desc(count)) %>% head(n=10)
bottoms <- counties.checklists %>% arrange(count) %>% head(n=10)
counties.checklists <- counties.checklists %>%
  mutate(top10 = ifelse(count %in% tops$count, TRUE, FALSE)) %>%
  mutate(bottom10 = ifelse(count %in% bottoms$count, TRUE, FALSE)) %>%
  arrange(desc(count))

# Pull out just the column of interest here ("count") and plot it.
tops2 <- counties.checklists %>%
  select(top10, geom)
ggplot() +
  geom_sf(data=tops2, aes(fill=top10)) +
  scale_fill_viridis_d(alpha = 1, option="D") +
  labs(title="Counties with large cities are heavily sampled",
       caption="Checklists from before 2021 are not included")
```

### Bottom 10 Counties
```{r bottom_county_table}
knitr::kable(head(arrange(by_county, count), n=10), caption="Bottom 10")
```

```{r bottom_county_map}
# Pull out just the column of interest here ("count") and plot it.
bottoms2 <- counties.checklists %>%
  select(bottom10, geom)
ggplot() +
  geom_sf(data=bottoms2, aes(fill=bottom10)) +
  scale_fill_viridis_d(alpha = 1, option="E") +
  labs(title="Counties with fewest checklists are in the NE Coastal Plains and piedmont",
       caption="Checklists from before 2021 are not included")
```

### Checklists per County Table
Note, the assignment of checklists to counties was performed by eBird and it is possible that errors exist due to traveling counts that cross county lines and errors in the recorded locations of checklists.
```{r per_county_all}
knitr::kable(by_county, caption="")
rm(by_county, tops, bottoms, tops2, bottoms2)
```

## Checklists per Atlas Block
Although coordinates are provided by eBird for checklists, they do not provide precise locations of eBirder effort for two reasons.  First, there are limits to the spatial precision of the points due to gps precision and/or observers ability to identify exactly where they birded on a map or in the app.  Second, many birders travel while birding but their paths are not available, only the distances they traveled.

Locational uncertainty is important and problematic because if it is large in relation to the level of analysis, it creates uncertainty about which spatial subregions, such as counties or atlas blocks, a checklist should be attributed to.  I explored two appraoches to handling locational uncertainty.

Method A -- Assign each checklist to the subregion that the checklist coordinate is located within.  This approach could generate deceptive results if checklists represent birding effort from multiple blocks but are assigned to a single block or if the coordinate is located in a block adjacent to where the birding actually occurred.  This approach should generally be expected to underestimate how many checklists covered some portion of a given block.

Method B -- Use polygons instead of the coordinates (points, Method A) in order to include the locational uncertainty.  Under this approach, coordinates are buffered with the distance traveled by the observer during the checklist period, plus 100 m to account for the fact that observers may have recorded birds at a distance from where they were located.  Each checklist is then assigned to all of the blocks that the polygon intersects in order to acknowledge that the checklist could represent effort from multiple blocks. Results from this approach can logically be expected to exaggerate the true footprint of birding effort and suggest blocks were sampled that actually were not, thus overestimating how many checklists covered some portion of a given block.  Furthermore, checklists with large effort distances produce enormous footprints than make results unhelpful.  Thus, I excluded checklists with effort distances greater than 5 km for this method.
```{r method_A}
method_A <- checklists %>%
  # Make checklists data frame spatial with right projection
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) %>%
  st_transform(6542) %>%
  # Find which blocks each coordinate is within
  st_join(blocks_sf, join = st_within, left=TRUE) %>%
  #################################select(-c(NAME, STATUS, COUNTY, FORM_LINK)) %>%
  # Summarize by number checklists within each block
  group_by(name) %>%
  summarize(checklists=n()) %>%
  select(name, checklists) %>%
  # Join back with blocks spatial frame to fill in zeros (as a data frame)
  data.frame() %>%
  select(-c(geometry)) %>%
  right_join(blocks_sf, by=("name" = "name")) %>%
  ##########################################select(name, checklists, geometry) %>%
  replace_na(list(checklists=0)) %>%
  # Make spatial again
  st_as_sf()
```

```{r buffered_coordinates}
# Create new sf of buffered checklist coordinates.  Some lists have no distance
#   so replace those with 0.  Stationary or short lists should be buffered 100
#   m to account for area surveyed.  This is imperfect, but more defensible than
#   nothing. Lists traveling > 5 km are just problematic and not informative so 
#   removed here.
buffered_coordinates <- checklists %>%
  # Make spatial frame
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) %>%
  st_transform(6542) %>%
  select(checklist_id, atlas_block, protocol_type, effort_distance_km,
         geometry) %>%
  # Buffer coordinates
  replace_na(list(effort_distance_km=0)) %>%
  filter(effort_distance_km <= 5) %>%
  mutate(buffer_length = (effort_distance_km + 0.1)*1000) %>%
  mutate(footprint = st_buffer(geometry, buffer_length)) %>%
  st_set_geometry("footprint") %>%
  select(-c(geometry)) 
```

```{r sample_buffered_coordinates}
# Create a list of blocks to sample from
sample_from <- c("Southport CE", "Southport CW", "Southport NE", "Southport NW",
                 "Southport SE", "Southport SW") 
                 
# quads <- c("SOUTHPORT", "FUNSTON", "WINNABOW", "LELAND", "CURRIE",
#            "SILK HOPE", "BYNUM", "WHITE CROSS", "EFLAND", "CEDAR GROVE",
#            "FRANKLIN", "ALARKA", "CLINGMANS DOME", "MT LECONTE", 
#            "BRYSON CITY")

sample_footprints <- checklists %>%
  # Make spatial frame
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) %>%
  st_transform(6542) %>%
  # Find which block each coordinate is within
  st_join(blocks_sf, join = st_within, left=TRUE) %>%
  # Use block name to pull out checklists assigned to blocks of interest
  filter(name %in% sample_from) %>%
  select(checklist_id, effort_distance_km, geometry) %>%
  # Buffer coordinates for footprints
  replace_na(list(effort_distance_km=0)) %>%
  filter(effort_distance_km <= 5) %>%
  mutate(buffer_length = (effort_distance_km + 0.1)*1000) %>%
  mutate(footprint = st_buffer(geometry, buffer_length)) %>%
  st_set_geometry("footprint") %>%
  select(-c(geometry)) 
```

```{r method_B}
# Intersect lists with blocks, tally checklists per block, drop footprint,
#   join with blocks sf to get block geometry
method_B <- buffered_coordinates %>%
  # Intersect footprints with blocks, NOTE this keeps "withins" and fragments
  st_intersection(blocks_sf) %>%
  # Find count by block
  group_by(name) %>%
  summarise(checklists = n()) %>%
  st_drop_geometry() %>%
  # Add zero blocks via a join
  right_join(blocks_sf, by="name") %>%
  select(name, checklists, geometry) %>%
  replace_na(list(checklists=0)) %>%
  st_as_sf()  
rm(buffered_coordinates)
```

```{r checklist_by_block_map}
# Get priority block centroids for plotting
priority_centroids <- st_centroid(filter(blocks_sf, priority==1)$geometry)

# Plot the map
tmap_mode("view")
tm_shape(method_A) +
  tm_fill(col = "checklists", palette = "Blues", style="fixed", alpha = .6,
          breaks = c(0, 1, 50, 100, 1000, 2000, 10000), as.count = TRUE) +
tm_shape(method_B) +
  tm_fill(col = "checklists", palette = "Reds", style="fixed", alpha = .6,
          breaks = c(0, 1, 50, 100, 1000, 2000, 10000), as.count = TRUE) +
tm_shape(priority_centroids) + tm_symbols(col="purple", size=0.01, shape=5) +
tm_shape(sample_footprints) + tm_borders()
```

Note: checklist footprints with center coordinates within select blocks are included in the map in order to display a sample of the underlying data for method B.  Priority centroids identify priority atlas blocks.

```{r checklist_by_block_cleanup}
rm(sample_footprints)
```

```{r checklists_per_block_boxplot}
# Join methods A and B results to facilitate plotting with facets
count_A <- method_A %>%
  st_drop_geometry() %>%
  select(name, checklists) %>%
  mutate(method = "A")
count_B <- method_B %>%
  st_drop_geometry() %>%
  select(name, checklists) %>%
  mutate(method = "B")
count_AB <- rbind(count_A, count_B)

# Boxplot
ggplot(data=count_AB) +
  geom_boxplot(mapping=aes(y=checklists, x=""), color="black", 
               outlier.colour="orange", show.legend=TRUE) + 
  coord_flip() + 
  labs(title="Many blocks have no checklists, but others have hundreds",
       subtitle="Method B tallies more checklists per block than method A",
       caption="Checklists from before 2021 are not included") +
  ylab("Checklists (n)") + 
  scale_y_continuous(n.breaks=12) +
  facet_wrap(~ method, nrow=2)

print("method A")
summary(count_A$checklists)
print("method B")
summary(count_B$checklists)
```


```{r blocks_with_most_checklists}
top_blocks_A <- count_A %>%
  arrange(desc(checklists)) %>%
  head(n = 20) %>%
  select(-c(method)) %>%
  rename(atlas_block=name)
knitr::kable(top_blocks_A, caption="Method A top 20 blocks")

top_blocks_B <- count_B %>%
  arrange(desc(checklists)) %>%
  head(n = 20) %>%
  select(-c(method)) %>%
  rename(atlas_block=name)
knitr::kable(top_blocks_B, caption="Method B top 20 blocks")
```

```{r checklists_priority_vs_no_chart}
count_AB$ysn_checklists <- ifelse(count_AB$checklists > 0, 
                                  "Checklists","No Checklists")

priority_blocks <- blocks_sf %>%
  data.frame() %>%
  filter(priority==1) %>%
  select(name, priority) %>%
  left_join(data.frame(count_AB), by=("name" = "name"))
  
ggplot(data=priority_blocks, aes(ysn_checklists)) +
  geom_bar() +
  ylab("Priority Blocks") +
  xlab("") + 
  labs(title="Several priority blocks remain unsampled",
       subtitle = "Method A suggests a higher number of unsampled priority blocks than method B",
       caption="Checklists from before 2021 are not included") +
  facet_wrap(~ method, ncol=2)
```

```{r checklist_block_count_cleanup}
rm(method_A, method_B, count_AB, count_A, count_B)
```


# Observers
## All observers
```{r observer_summary_stats}
by_observer <- checklists %>%
  group_by(observer_id) %>%
  summarize(
    count = n())
obs_summary <- by_observer$count %>%
  summary()
```

Most eBirders have only submitted a handful of checklists. The median number of checklists submitted is only `r obs_summary[[3]]`, and observers who have submitted more than `r obs_summary[[5]]` are in the top quartile for number of submissions.
```{r observer_boxplot}
ggplot(data=by_observer) +
  geom_boxplot(mapping=aes(y=count, x=""), color="darkgreen", 
               outlier.colour="blue", show.legend=TRUE) + 
  coord_flip() + 
  labs(title="Even moderately active eBirders are in the minority",
       caption="Checklists from before 2021 are not included") +
  ylab("Checklists (n)") + 
  scale_y_continuous(n.breaks=12)
```

## Core observers
It is informative to look more closely at a subset of active observers in order to better characterize observers who may participate in the atlas.  Here, I exclude the extreme outliers and people with fewer than some arbitrarily "small" number of submissions (20). 
```{r core_observer_stats}
core_observers <- by_observer %>%
  filter(count > 20 & count < 3000)
topobs_summary <- core_observers$count %>%
  summary()
```

```{r core_observer_boxplot}
ggplot(data=core_observers) +
  geom_boxplot(mapping=aes(y=count, x=""), color="darkblue", 
               outlier.colour="orange", show.legend=TRUE) + 
  coord_flip() + 
  labs(title='Most "core" eBirders have submitted < 100 checklists',
       caption="Checklists from before 2021 were not included in this summary") +
  ylab("Checklists (n)") + 
  scale_y_continuous(n.breaks=12)
```

# Protocol
```{r protocol}
# Summarize protocol
by_protocol <- checklists %>%
  group_by(protocol_type) %>%
  summarize(
    count = n())
colnames(by_protocol) = c("Protocol Type", "Checklists (n)")

# Print table
knitr::kable(by_protocol)
```

# Project
```{r project}
# Summarize projects
by_project <- checklists %>%
  group_by(project_code) %>%
  summarize(
    count = n())
colnames(by_project) = c("Project", "Checklists (n)")

# Print table
knitr::kable(by_project)
```

# Start Time
```{r time}
start_df <- checklists %>%
  select(time_observations_started) %>%
  mutate(time=hour(as_hms(time_observations_started)))

ggplot(data=start_df) +
  geom_boxplot(mapping=aes(y=time, x=""), color="darkblue", 
               outlier.colour="magenta", show.legend=TRUE) + 
  coord_flip() + 
  labs(title="eBirders generally get out in the late morning",
       caption="Checklists from before 2021 are not included") +
  ylab("Start Hour") + 
  scale_y_continuous(n.breaks=12)
```

# Duration
```{r duration_prep}
# Prep a data frame of duration data
duration_df <- checklists %>%
  select(duration_minutes) %>% 
  arrange(desc(duration_minutes))

too_long <- subset(duration_df, duration_minutes > 1440)
sorta_long <- subset(duration_df, duration_minutes <1441 & duration_minutes > 720)
duration_OK <- subset(duration_df, duration_minutes < 721)
```

* The durations of most checklists are less than one day (1,440 minutes), but `r nrow(too_long)` checklists carried on longer than a day and `r nrow(sorta_long)` checklists lasted between 12 and 24 hours. 
* I omitted checklists longer than 12 hours for the summaries in this section.

```{r duration_boxplot}
ggplot(data=duration_OK) +
  geom_boxplot(mapping=aes(y=duration_minutes, x=""), color="darkblue", 
               outlier.colour="orange", show.legend=TRUE) + 
  coord_flip() + 
  labs(title="Many checklists were over an hour long",
       caption="Checklists from before 2016 are not included") +
  ylab("Minutes") + 
  scale_y_continuous(n.breaks=12)
```

```{r duration_table}
print(summary(duration_OK$duration_minutes))
```

# Distance
```{r distance_prep}
# Prep a data frame of duration data
distance_df <- checklists %>%
  select(effort_distance_km) %>% 
  arrange(desc(effort_distance_km))

too_far <- subset(distance_df, effort_distance_km > 5)
distance_OK <- subset(distance_df, effort_distance_km <= 5)
```

```{r distance_boxplot}
ggplot(data=distance_OK) +
  geom_boxplot(mapping=aes(y=effort_distance_km, x=""), color="darkgreen", 
               outlier.colour="blue", show.legend=TRUE) + 
  coord_flip() + 
  labs(title="Checklists with distances < 1km are in the minority",
       caption="Checklists from before 2021 are not included") +
  ylab("Kilometers") + 
  scale_y_continuous(n.breaks=12)
summary(distance_OK$effort_distance_km)
```

```{r too_far_boxplot}
ggplot(data=too_far) +
  geom_boxplot(mapping=aes(y=effort_distance_km, x=""), color="darkblue", 
               outlier.colour="orange", show.legend=TRUE) + 
  coord_flip() + 
  labs(title="Most excessively long checklists are around 8km in length",
       caption="Checklists from before 2021 are not included") +
  ylab("Kilometers") + 
  scale_y_continuous(n.breaks=12)
summary(too_far$effort_distance_km)
```

# Runtime
```{r runtime}
endtime <- Sys.time()
runtime <- endtime - starttime
```
`r format(runtime)`