---
title: "Code-centered Behavior Code Review"
author: "N.M. Tarr"
date: "2023-12-04"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
if(!require(auk)) install.packages(
  "auk", repos = "http://cran.us.r-project.org")
if(!require(tmap)) install.packages(
  "tmap", repos = "http://cran.us.r-project.org")
if(!require(here)) install.packages(
  "here", repos = "http://cran.us.r-project.org")

# print(file.path("C:","Users", "skanderson","OneDrive - State of North Carolina","@@ncba","ncba"))
# setwd("C://Users//skanderson//OneDrive - State of North Carolina//@@ncba//ncba")
# here::i_am(path=file.path("C:","Users", "skanderson","OneDrive - State of North Carolina","@@ncba","ncba","readme.md"))

source(here("resources","ncba_functions.R"))

```

# Purpose
At times it may be advantageous to review all records that have a certain behavior code, regardless of the species.  This document demonstrates how NCBA code can facilitate that, as well as for review of groups of species.

# All Species Review
The "PE" breeding code is for observations of birds with physiological conditions that indicate breeding.  It should predominantly justified by in-hand observations of birds.  It is not often used and it may be worth while to review all cases where it has been used.

The "get_breeding_records" function can retrieve records of "PE".
```{r}
pe <- get_breeding_records(behaviors = c("PE")) %>%
  to_EBD_format() %>%
  auk_unique() %>%
  filter(protocol_type != "Banding") %>%
  arrange(common_name)

print(select(pe, c(sampling_event_identifier, common_name, protocol_type, has_media)))
write_csv(pe, paste(work_dir, "physiological_evidence.csv"))
```

See for which species the code was reported.
```{r}
print(unique(pe$common_name))
```

The records can be mapped in a way where clicking on the records symbol provides a hyperlink to the checklist's website.
```{r}
# Make data frames of suspicious records and their uncertainty buffers
pe_sf <- records_as_sf(pe, kind = "observations", 
                            method = "points") %>%
  right_join(pe, by = "sampling_event_identifier")

uncertainty_buffer <- records_as_sf(pe, kind = "observations", 
                              method = "point-radius") %>%
  filter(buffer_length > 0)

# Add a column with code to open the webpage for each checklist
pe_sf$front <- '<a href = https://ebird.org/checklist/'

# Strip off any excess identifiers; group checklists produce "S104604778,S104604779"
pe_sf$sampling_event_identifier <- lapply(strsplit(pe_sf$sampling_event_identifier, split = ","), function(l) l[[1]])
pe_sf$URL <- with(pe_sf, paste0(front, sampling_event_identifier,
                                            ">visit</a>"))

# Draw the map
tmap_mode("view") 
tm_shape(counties_NC(), name = "Counties") + tm_borders() +
tm_shape(pe_sf, name = "PE records") + 
  tm_dots(popup.vars = c("URL", "common_name"), popup.format = list(html.escape = F)) + 
tm_shape(uncertainty_buffer, name = "Locational Uncertainty") + tm_borders(col = "green") + 
  tm_layout(title = "PE records")
```

The records can also be summarized and plotted by day of year with a boxplot that also provides hyperlinks for each record.
```{r}
breeding_boxplot(species = NULL, data = pe, type = "interactive")
```


# Species Group
The process from above can be applied to a group of species with inclusion of a statement that filters based on a species list.  For this example, I present records of tanagers exhibiting distraction displays.

The "get_breeding_records" function can retrieve records of "A".
```{r}
species <- c("Summer Tanager", "Scarlet Tanager")

df <- get_breeding_records(behaviors = c("A")) %>%
  to_EBD_format() %>%
  auk_unique() %>%
  filter(common_name %in% species)

print(df %>% select(sampling_event_identifier, common_name, behavior_code))
```

The records can be mapped in a way where clicking on the records symbol provides a hyperlink to the checklist's website.
```{r}
# Make data frames of suspicious records and their uncertainty buffers
A_sf <- records_as_sf(df, kind = "observations", 
                            method = "points")

uncertainty_buffer <- records_as_sf(df, kind = "observations", 
                              method = "point-radius") %>%
  filter(buffer_length > 0)

# Add a column with code to open the webpage for each checklist
A_sf$front <- '<a href = https://ebird.org/checklist/'

# Strip off any excess identifiers; group checklists produce "S104604778,S104604779"
A_sf$sampling_event_identifier <- lapply(strsplit(A_sf$sampling_event_identifier, split = ","), function(l) l[[1]])
A_sf$URL <- with(A_sf, paste0(front, sampling_event_identifier,
                                            ">visit</a>"))

# Draw the map
tmap_mode("view") 
tm_shape(counties_NC(), name = "Counties") + tm_borders() +
tm_shape(A_sf, name = "FS records") + 
  tm_dots(popup.vars = c("URL"), popup.format = list(html.escape = F)) + 
tm_shape(uncertainty_buffer, name = "Locational Uncertainty") + tm_borders(col = "green")
```

The records can also be summarized and plotted by day of year with a boxplot that also provides hyperlinks for each record.
```{r}
breeding_boxplot(species = NULL, data = df, type = "interactive")
```