---
title: "demo_user_stats"
author: "N.M. Tarr"
date: "2023-10-05"
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```

Identify an observer.
```{r}
observer <- "obsr1000095"
```

Get all of their checklists.
```{r}
checklists.obs <- get_checklists(observer = observer, project = NULL) %>%
  to_EBD_format()
```

View the checklist records.  They could also be saved with to.csv().
```{r}
View(checklists.obs)
```

Plot the checklist locations as points.
```{r}
sf <- records_as_sf(checklists.obs, kind = "checklists", method = "points")

# Make a simple map
ggplot(data = counties_NC()) +
    geom_sf() +
    geom_sf(data = sf, size = 4, shape = 20, 
            crs = st_crs(6542), colour = "darkred") + 
    ggtitle(paste("Coordinates of Checklists from", observer), subtitle = "")
```

Plot the checklist locations as uncertainty buffers.
```{r}
sf <- records_as_sf(checklists.obs, kind = "checklists", method = "point-radius")

# Make a simple map
ggplot(data = counties_NC()) +
    geom_sf() +
    geom_sf(data = sf, size = 4, shape = 20, 
            crs = st_crs(6542), colour = "darkred") + 
    ggtitle(paste("Uncertainty Buffers of Checklists from", observer), 
                  subtitle = "")
```

Summarise the protocol type of the checklists.
```{r}
protocol_table(checklists.obs)
show(protocol_type_pie(checklists.obs))
```

Summarize the breakdown of complete vs. incomplete checklists.
```{r}
complete_checklist_table(checklists.obs)
```

Summarize the distance and duration of checklists.
```{r}
duration_distance_table(checklists.obs)
show(effort_distance_boxplot(checklists.obs))
show(duration_minutes_boxplot(checklists.obs))
```

Get all the observer's observations.
```{r}
observations <- get_observations(observer = observer, project = NULL) %>%
  to_EBD_format()
```







