---
title: "Exploration of Data for an Individual Species"
author: "N.M. Tarr"
date: "2023-10-9"
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# setwd("C:/Users/skanderson/OneDrive - State of North Carolina/@@ncba/ncba/resources")
source("ncba_functions.R")
library(auk)

knitr::opts_knit$set(root.dir = work_dir) 
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# Purpose
This document demonstrates species-level data summaries that are available from the NCBA code repository.

Identify a species and get the observations.
```{r}
species <- "Grasshopper Sparrow"

observations <- get_observations(species = species, project = "EBIRD_ATL_NC",
                                 EBD_fields_only = FALSE) %>%
  to_EBD_format() %>%
  auk_unique()
```

Print the number of observations.
```{r}
print(nrow(observations))
```

Print the total number of individuals reported.
```{r}
print(sum(as.integer(observations$observation_count)))
```


Plot the coordinates.
```{r}
# Get spatial data frame of observations
obs.spatial <- records_as_sf(records_df = observations, 
                             kind = "observations", method = "points")

# Make a simple map
ggplot(data = counties_NC()) +
    geom_sf() +
    geom_sf(data = obs.spatial, size = 4, shape = 20, 
            crs = st_crs(6542), colour = "darkred", fill = NA) + 
    ggtitle(paste(species, "Record Coordinates"), subtitle = "")
```

Plot the footprints.
```{r}
# Get spatial data frame of observations
obs.spatial <- records_as_sf(records_df = observations, 
                             kind = "observations", method = "point-radius")

# Make a simple map
ggplot(data = counties_NC()) +
    geom_sf() +
    geom_sf(data = obs.spatial, size = 4, shape = 20, 
            crs = st_crs(6542), colour = "darkred", fill = NA) + 
    ggtitle(paste(species, "Uncertainty Buffers"), subtitle = "")
```

Map the highest reported breeding code by block with needed blocks identified.
```{r}
show(map_needed_highest(species = species))
```

Get a boxplot of breeding phenology
```{r}
breeding_boxplot(species = species, data = observations, type = "ecoregional", 
                 drop = FALSE, lump = breeding_codes())
```

Plot observations by week.
```{r}
lists_by_week(observations)
```

Summarize checklist distance.
```{r}
show(effort_distance_boxplot(observations))
```

Summarize duration
```{r}
show(duration_minutes_boxplot(observations))
```

```{r}
duration_distance_table(observations)
```

Summarize start times.
```{r}
show(start_time_boxplot(observations))
```

Summarize locality type.
```{r}
show(locality_type_pie(observations))
```

Summarize protocol in a table.
```{r}
protocol_table(observations)
```

Summarize protocol in a figure.
```{r}
protocol_table(observations)
```

Summarize checklist completeness.
```{r}
complete_checklist_table(observations)
```

Summarize observations per year.
```{r}
show(year_bar(observations))
```

