---
title: "demo_block_stats"
author: "N.M. Tarr"
date: "2023-10-10"
output: word_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/Code/NCBA/resources")
source("ncba_functions.R")
library(auk)

knitr::opts_knit$set(root.dir = work_dir)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE)
```


# Purpose
This document details and demonstrates code that provides information about the data available for an individual block.  Questions of block status will likely be season-specific, so this demo is organized by season.

First, specify a block of interest.
```{r}
block <- "35079H1SE"
```

Get identifying information about the block
```{r}
blocks <- get_blocks(spatial = TRUE)
block_info <- blocks %>% 
  filter(ID_BLOCK_CODE == block) %>%
  select(c("ID_BLOCK_CODE", "ID_BLOCK", "ID_EBD_NAME", "ECOREGION"
  ))
print(block_info %>% data.frame())
```

Highlight the block on a map
```{r}
# Make a simple map
ggplot() +
    geom_sf(data = blocks, fill = "lightgrey", color = "darkgrey") +
    geom_sf(data = block_info, fill = "purple", color = "purple") +
    ggtitle("", subtitle = "")
```

Get a list of species predicted by National GAP to occur in the block in each season.
```{r}
preds <- block_predicted_spp(block = block, source = "GAP")
predicted_summer <- preds$summer
predicted_winter <- preds$winter
```


## Summer/Breeding Season
Breeding season dates vary by species, but we need to choose a single start and end date.  Therefore, look at some calculated start and end dates (excluding outliers) for common, early, and late breeders to inform a choice in dates.  The flock of interest is in the Piedmont, so focus on Piedmont dates.

The Northern Cardinal seems to breed as soon as possible given weather and is very common.
```{r}
print(calculate_breeding_dates(species = "Northern Cardinal",
                               basis = c("confirmed", "probable"),
                               quantiles = c(0.05, 0.95),
                               year = 2023,
                               year_day = FALSE))
```

The American Goldfinch is a later breeder.
```{r}
print(calculate_breeding_dates(species = "American Goldfinch",
                               basis = c("confirmed", "probable"),
                               quantiles = c(0.05, 0.95),
                               year = 2023,
                               year_day = FALSE))
```

The Carolina Wren seems to breed several times per year if possible.
```{r}
print(calculate_breeding_dates(species = "Carolina Wren",
                               basis = c("confirmed", "probable"),
                               quantiles = c(0.05, 0.95),
                               year = 2023,
                               year_day = FALSE))
```

Based on the results of the breeding date calculations, it seems that March 11 - August 18 is a good choice for the Piedmont.  Get species lists from the Atlas Cache for the time period between those days.
```{r}
start_day <- 70
end_day <- 230
observed_summer <- block_spp_lists(block = block, start_day = start_day, 
                                   end_day = end_day, within = TRUE)
```

Which species were predicted to occur during summer but have not been observed?
```{r}
print(setdiff(predicted_summer, observed_summer$all))
```

Which species were not predicted to occur but were then observed?
```{r}
print(setdiff(observed_summer$all, predicted_summer))
```

Which species have been observed, but not coded as probable or confirmed during summer?
```{r}
print(setdiff(observed_summer$all, union(observed_summer$probable, observed_summer$confirmed)))
```

Get a summary of how many species have been submitted for each breeding category.
```{r}
print(spp_count_summary(observed_summer, predicted_summer))
```

Summarize the duration of effort in hours.
```{r}
print(summarize_duration(block = block, summarize_by = "year", start_day = start_day,
                         end_day = end_day, within = TRUE))
```


## Winter
Assign start and end days to for winter.  The concept of winter regarding birds is more ambiguous than summer, so use the NCBA winter start and end days.
```{r}
start_day <- 59
end_day <- 305
```

Get species lists from the Atlas Cache for the time period between those days.
```{r}
observed_winter <- block_spp_lists(block = block, start_day = start_day, 
                                   end_day = end_day, within = FALSE)
```

Which species were predicted to occur during winter but have not been observed?
```{r}
print(setdiff(predicted_winter, observed_winter$all))
```

Which species were not predicted to occur but were then observed?
```{r}
print(setdiff(observed_winter$all, predicted_winter))
```

Get a summary of how many species have been submitted for each breeding category.  Reports of breeding during the winter dates reflects upon the quality or appropriateness of the records and/ or the winter start and end dates.
```{r}
print(spp_count_summary(observed_winter, predicted_winter))
```

Which species have been confirmed breeding during winter?
```{r}
print(observed_winter$confirmed)
```

Which species have been reported possibly breeding during winter?
```{r}
print(observed_winter$possible)
```

Which species have been reported probably breeding during winter?
```{r}
print(observed_winter$probable)
```

Summarize duration of effort in hours by month.
```{r}
print(summarize_duration(block = block, summarize_by = "month", start_day = start_day,
                         end_day = end_day, within = FALSE))
```